### 2025-01-23 scripts/compare-hard-shell-with-accessory.Rmd
add first results files

# Compare hard shell genes vs. accessory

```{r}
library(ggplot2)
```

```{r}
# Pangenome statistics for PTUs
ptu = read.csv('../data/PTU-pangenome-stats.csv')

# Read in RMES comparisons
df_4 = read.csv('~/Downloads/trieste/test_k4.csv')
df_6 = read.csv('~/Downloads/trieste/test_k6.csv')
df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
```

Do the statistical test

```{r}
wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE)

```
```{r}
ggplot(df_4_filter, aes(x="k=4", wilcoxon_stat, colour=p_value<0.05))+
  geom_point()
```

### 2025-01-23 scripts/compare-hard-shell-with-accessory.R

@@ -1,106 +1,0 @@
-# Compare hard shell genes vs. accessory
-library(ggplot2)
-
-# Location of hard shell (core) and accessory fasta files
-FASTA_DIR = '~/Downloads/trieste/'
-
-# Read in PTU stats and only keep those with >9 hard shell genes
-ptu.stats = read.csv('../data/PTU-pangenome-stats.csv', 
-                     header=T)
-SUBSET_PTUS = ptu.stats$PTU[ptu.stats$pangenome.category=="include"]
-  
-rm_target_db = read.csv('../data/RM_target_db.csv')
-
-compareHardShellVsAccessory = function(PTU_name, k){
-  ptu_df_hard_shell = read.csv(paste0(FASTA_DIR, 
-                                      PTU_name,
-                                      '_core_genes_k', k, '.csv'), 
-                               header=T)
-    ptu_df_accessory = read.csv(paste0(FASTA_DIR, 
-                                     PTU_name,
-                                     '_accessory_genes_k', k, '.csv'), 
-                              header=T)
-  # Normalise scores by maximum (for easier comparison)
-  ptu_df_hard_shell$score.norm = ptu_df_hard_shell$score/max(ptu_df_hard_shell$score)
-  ptu_df_accessory$score.norm = ptu_df_accessory$score/max(ptu_df_accessory$score)
-  # There would be a better scaling here...
-  
-  ptu_df = merge(ptu_df_hard_shell, ptu_df_accessory, by="word")
-  
-  palindromes = tolower(read.csv(paste0('~/Dropbox/_Projects/2023-Trieste/2023-Trieste-RM/palindromes-k', k, '.txt'),
-                                 header=F)$V1)
-  ptu_df$palindrome = ifelse(ptu_df$word %in% palindromes, "palindrome", "non.palindrome")
-  ptu_df$score.discrepancy = ptu_df$score.x-ptu_df$score.y
-  
-  # ggplot(ptu_df, aes(score.norm.x, score.norm.y))+
-  #   geom_point()+
-  #   theme_minimal()+
-  #   facet_wrap(~palindrome)
-  p = ggplot(ptu_df, aes(palindrome, score.discrepancy))+
-    ggbeeswarm::geom_quasirandom()+
-    #geom_boxplot()+
-    ggsignif::geom_signif(comparisons = list(c("palindrome", "non.palindrome")),
-                          test = "wilcox.test")+
[…]

### 2025-01-23 scripts/deambiguate-target-db.py
deambiguate target DB

@@ -1,0 +1,62 @@
+import pandas as pd
+
+ambiguity_codes =   {'A': ['A'],\
+                    'G' : ['G'],\
+                    'C'	: ['C'],\
+                    'T' : ['T'],\
+                    'Y'	: ['C', 'T'],\
+                    'R'	: ['A','G'],\
+                    'W'	: ['A','T'],\
+                    'S'	: ['G','C'],\
+                    'K'	: ['T','G'],\
+                    'M'	: ['C','A'],\
+                    'D'	: ['A','G','T'],\
+                    'V'	: ['A','C','G'],\
+                    'H'	: ['A','C','T'],\
+                    'B'	: ['C','G','T'],\
+                    'X'	: ['A','C','G','T'],\
+                    'N'	: ['A','C','G','T'],\
+                    '-' : ['-']}
+
+def possibleSequences(dna_sequence):
+    '''Takes a DNA sequence (possibly containing ambiguous bases) and returns
+    a list of all possible sequences.
+    Args:
+        dna_sequence (str)
+            String of DNA sequence (uppercase)
+    Returns:
+        possible_strings (list)
+            List of strings - all possible DNA sequences
+    '''
+    # Get all possible bases at each position
+    possible_bases = [ambiguity_codes[base] for base in dna_sequence]
+    # Go through and add these on
+    possible_strings = []
+    for bases in possible_bases:
+        if len(possible_strings)==0: # if no strings yet, use first base
+            possible_strings = list(bases)
+        elif len(bases)==1: # if just one possible base at position, add it on
+            possible_strings = [x+bases[0] for x in possible_strings]
+        else: # if multiple possibilities, add them on one by one
+            additions = []
+            for base in bases:
+                additions.append([x+base for x in possible_strings])
+            possible_strings = [x for addition in additions for x in addition]
+    return possible_strings
+
+
+target_db = pd.read_csv("data/RM_target_db.csv")
+new_target_db = []
+for i, sequence in enumerate(target_db['sequence']):
+    new_sequences = possibleSequences(sequence)
+    row = list(target_db.loc[i])
+    for s in new_sequences:
+        new_target_db.append(row+[s])
+
+new_target_db_df = pd.DataFrame(new_target_db)
+new_target_db_df.columns = ['genus', 'sequence', 'genome', 'total_genomes', 'normalized_count', 'deambiguated_sequence']
+new_target_db_df.to_csv("data/RM_target_db_deambiguated.csv", index=False)
+
+
+
+

### 2025-01-23 scripts/deambiguate-target-db.py
add length to db

@@ -55,6 +55,7 @@ for i, sequence in enumerate(target_db['sequence']):
 
 new_target_db_df = pd.DataFrame(new_target_db)
 new_target_db_df.columns = ['genus', 'sequence', 'genome', 'total_genomes', 'normalized_count', 'deambiguated_sequence']
+new_target_db_df = new_target_db_df.assign(length=[len(x) for x in new_target_db_df['deambiguated_sequence']])
 new_target_db_df.to_csv("data/RM_target_db_deambiguated.csv", index=False)
 
 

### 2025-01-24 scripts/create-PTU-target-DB.py
more notes

@@ -1,0 +1,93 @@
+import csv
+import os
+
+
+def read_db_to_dict(file_path, key_column, value_column, delimiter=','):
+    """
+    Reads a file and returns list of unique values in value_column for each unique value of key_column
+    Args:
+        file_path (str): Path to the input file.
+        key_column (str): Name of the column to use as the key.
+        value_column (str): Name of the column to use as the value.
+        delimiter (str): Delimiter for the file (default is tab-delimited).
+        
+    Returns:
+        dict: A dictionary where keys are unique values from key_column and
+              values are lists of unique items from value_column.
+    """
+    summary_dict = {}
+    
+    # Open and read the file
+    with open(file_path, 'r') as file:
+        reader = csv.DictReader(file, delimiter=delimiter)
+        
+        # Populate the dictionary
+        for row in reader:
+            key = row[key_column]
+            value = row[value_column]
+            if key not in summary_dict:
+                summary_dict[key] = []
+            if value not in summary_dict[key]:  # Avoid duplicates
+                summary_dict[key].append(value)
+    
+    return summary_dict
+
+def combine_target_dicts(targets_by_genera, genera_by_PTU):
+    """
+    Combines the dictionaries to create a new dictionary of targets by PTU.
+    Args:
+        targets_by_genera (dict): Dictionary with {genus: [list_of_targets]}.
+        genera_by_PTU (dict): Dictionary with {PTU: [list_of_genera]}.
+    
+    Returns:
+        dict: A dictionary with {PTU: [list_of_targets]}.
+    """
+    targets_by_PTU = {}
+    
+    for ptu, genera in genera_by_PTU.items():
+        # Initialize an empty list for the current PTU
+        targets_by_PTU[ptu] = []
+        
+        # Collect all targets for the genera associated with the current PTU
+        for genus in genera:
+            if genus in targets_by_genera:
+                targets_by_PTU[ptu].extend(targets_by_genera[genus])
+        
+        # Remove duplicates from the list of targets
+        targets_by_PTU[ptu] = list(set(targets_by_PTU[ptu]))
+    
+    return targets_by_PTU
+
+
+
+def write_target_files(targets_by_PTU, output_directory):
+    """
+    Writes a file for each PTU with its list of targets.
+    
+    Args:
+        targets_by_PTU (dict): Dictionary with {PTU: [list_of_targets]}.
[…]

### 2025-01-24 scripts/rmes-score-discrepancy.py
more flexible scripts to allow passing different kmer targets per PTU

@@ -49,6 +49,7 @@ def calculate_discrepancy_and_test(core_file, accessory_file, kmers_file):
         wilcox_stat, p_value = None, None
 
     return {
+        'n_special_kmers' : len(kmers_data),
         'median_discrepancy_kmers': median_discrepancy_kmers,
         'median_discrepancy_others': median_discrepancy_others,
         'wilcoxon_statistic': wilcox_stat,
@@ -61,5 +62,5 @@ if __name__ == "__main__":
     results = calculate_discrepancy_and_test(args.core, args.accessory, args.kmers)
     filename_base = args.core.split('_')[0].rsplit("/", 1)[-1]
 
-    results = [filename_base,results['median_discrepancy_kmers'], results['median_discrepancy_others'], results['wilcoxon_statistic'],results['p_value']]
+    results = [filename_base,results['n_special_kmers'], results['median_discrepancy_kmers'], results['median_discrepancy_others'], results['wilcoxon_statistic'],results['p_value']]
     print(','.join([str(x) for x in results]))

### 2025-01-24 scripts/run-rmes-score-discrepancy.py

@@ -6,15 +6,16 @@ def run_comparisons(base_script, directory, kmers_file, output_file, k):
     # Find all core and accessory files in the given directory
     core_files = sorted(glob.glob(os.path.join(directory, "PTU-*_core_genes_k"+str(k)+".csv")))
     accessory_files = sorted(glob.glob(os.path.join(directory, "PTU-*_accessory_genes_k"+str(k)+".csv")))
-
+    if kmers_file=="per-PTU":
+        kmers_files = sorted(glob.glob(os.path.join(directory, "PTU-*_targets.txt")))
     # Ensure matching pairs
     if len(core_files) != len(accessory_files):
         print("Error: Mismatched number of core and accessory files.")
         return
 
     with open(output_file, "w") as out:
-        out.write("PTU,discrepancy_selected,discrepancy_others,wilcoxon_stat,p_value\n")
-        for core_file, accessory_file in zip(core_files, accessory_files):
+        out.write("PTU,n_selected_kmers,discrepancy_selected,discrepancy_others,wilcoxon_stat,p_value\n")
+        for core_file, accessory_file, kmers_file in zip(core_files, accessory_files, kmers_files):
                 # Construct the command
                 command = [
                     "python", base_script,
@@ -40,7 +41,7 @@ if __name__ == "__main__":
     parser = argparse.ArgumentParser(description="Run comparisons for multiple core and accessory files.")
     parser.add_argument("--script", required=True, help="Path to the comparison script.")
     parser.add_argument("--directory", required=True, help="Directory containing core and accessory files.")
-    parser.add_argument("--kmers", required=True, help="Path to the kmers text file (make sure it matches value of k).")
+    parser.add_argument("--kmers", required=True, help="Path to the kmers text file (make sure it matches value of k). If 'per-PTU' is passed then will use targets by PTU, generated by scripts/create-PTU-target-DB.py")
     parser.add_argument("--output", required=True, help="Path to the output file.")
     parser.add_argument("--k", required=True, help="Value of k (4,5,6).")
 

### 2025-01-24 scripts/compare-hard-shell-with-accessory.Rmd
deambiguate target DB
analysis of RM targets

@@
 library(ggplot2)
+library(knitr)
+library(xtable)
 ```
@@
 ```{r}
+wilcox.test(df_4_filter$discrepancy_selected, df_4_filter$discrepancy_others, paired = TRUE)
 wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE)
@@
 ```
+
+```{r }
+df_4 = read.csv('~/Downloads/trieste/test_4mer_targets.csv')
+df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
+df_5 = read.csv('~/Downloads/trieste/test_5mer_targets.csv')
+df_5_filter = df_5[df_5$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
+df_6 = read.csv('~/Downloads/trieste/test_6mer_targets.csv')
+df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
+
+wilcox_tests = list(wilcox.test(df_4_filter$discrepancy_selected, df_4_filter$discrepancy_others, paired = TRUE),
+                    wilcox.test(df_5_filter$discrepancy_selected, df_5_filter$discrepancy_others, paired = TRUE),
+                    wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE))
+results_df = data.frame(median_discrepancy_targets=c(median(df_4_filter$discrepancy_selected),
+                                        median(df_5_filter$discrepancy_selected),
+                                        median(df_6_filter$discrepancy_selected)),
+           median_discrepancy_others=c(median(df_4_filter$discrepancy_others),
+                                        median(df_5_filter$discrepancy_others),
+                                        median(df_6_filter$discrepancy_others)),
+           wilcoxon_stat=c(wilcox_tests[[1]]$statistic,
+                           wilcox_tests[[2]]$statistic,
+                           wilcox_tests[[3]]$statistic),
+           p_value=c(wilcox_tests[[1]]$p.value,
+                     wilcox_tests[[2]]$p.value,
+                     wilcox_tests[[3]]$p.value))
+kable(results_df, digits = 3, format = "html")
+
+```
+
+```{r}
+# Read in results for RM targets
+df_4 = read.csv('~/Downloads/trieste/test_RM_targets_k4.csv')
+df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_4$discrepancy_selected!="None"],]
+df_5 = read.csv('~/Downloads/trieste/test_RM_targets_k5.csv')
+df_5_filter = df_5[df_5$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_5$discrepancy_selected!="None"],]
+df_6 = read.csv('~/Downloads/trieste/test_RM_targets_k6.csv')
+df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_6$discrepancy_selected!="None"],]
+
+# Convert discrepancies to numeric (presence of None has messed this up)
+df_4_filter$discrepancy_selected = as.numeric(df_4_filter$discrepancy_selected)
+df_5_filter$discrepancy_selected = as.numeric(df_5_filter$discrepancy_selected)
+df_6_filter$discrepancy_selected = as.numeric(df_6_filter$discrepancy_selected)
+
+wilcox_tests = list(wilcox.test(df_4_filter$discrepancy_selected, df_4_filter$discrepancy_others, paired = TRUE),
+                    wilcox.test(df_5_filter$discrepancy_selected, df_5_filter$discrepancy_others, paired = TRUE),
+                    wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE))
+results_df = data.frame(n_PTUs = c(nrow(df_4_filter),
+                                   nrow(df_5_filter),
+                                   nrow(df_6_filter)),
+                        median_n_targets = c(median(df_4_filter$n_selected_kmers),
+                                   median(df_5_filter$n_selected_kmers),
+                                   median(df_6_filter$n_selected_kmers)),
+  median_discrepancy_targets=c(median(df_4_filter$discrepancy_selected),
+                                        median(df_5_filter$discrepancy_selected),
+                                        median(df_6_filter$discrepancy_selected)),
+           median_discrepancy_others=c(median(df_4_filter$discrepancy_others),
+                                        median(df_5_filter$discrepancy_others),
+                                        median(df_6_filter$discrepancy_others)),
+           wilcoxon_stat=c(wilcox_tests[[1]]$statistic,
+                           wilcox_tests[[2]]$statistic,
+                           wilcox_tests[[3]]$statistic),
+           p_value=c(wilcox_tests[[1]]$p.value,
+                     wilcox_tests[[2]]$p.value,
+                     wilcox_tests[[3]]$p.value))
+kable(results_df, digits = 3, format = "html")
+```

### 2025-01-24 scripts/compare-hard-shell-with-accessory.Rmd
new results files with discrepancies for within-range targets

@@
 # Read in results for RM targets
-df_4 = read.csv('~/Downloads/trieste/test_RM_targets_k4.csv')
+df_4 = read.csv('../results/rmes_discrepancies_targets_k4.csv')
 df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_4$discrepancy_selected!="None"],]
-df_5 = read.csv('~/Downloads/trieste/test_RM_targets_k5.csv')
+df_5 = read.csv('../results/rmes_discrepancies_targets_k5.csv')
 df_5_filter = df_5[df_5$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_5$discrepancy_selected!="None"],]
-df_6 = read.csv('~/Downloads/trieste/test_RM_targets_k6.csv')
+df_6 = read.csv('../results/rmes_discrepancies_targets_k6.csv')
 df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_6$discrepancy_selected!="None"],]

### 2025-01-27 scripts/Paper-analysis.Rmd
one Rmarkdown for paper analysis

--- 2025-01-24 scripts/compare-hard-shell-with-accessory.Rmd
@@
-# Compare hard shell genes vs. accessory
-
 ```{r}
@@
 ```
+
+## Compare hard shell genes vs. accessory
+
+### Short palindromes (k=4, 6)
 
@@
 ```
+
+### All RM targets (not included in main text)
 
@@
 ```
+
+### Hard shell vs. accessory for within-range RM targets
 
@@
 ```
+
+## Analysis of leading region
+
+Some preliminary statistics about discarded PTUs (for main text).
+
+```{r}
+lr = read.csv('data/leading_region_081123.tsv', sep='\t')
+plasmid.table = read.csv('data/plasmid_table.tsv', sep='\t')
+retained.PTUs.lr = read.csv('data/leading-region-PTUs.txt', header=F)$V1
+lr.identified.PTU.names = sort(unique(lr$PTU))
+# Minimum sizes
+sort(sapply(lr.identified.PTU.names, function(x) 
+  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
+```
+
+We discard those where the minimum size is <20kb: E1, E3, E9, E10, E14, E76. We check the maximum size per PTU:
+
+```{r}
+sort(sapply(paste0("PTU-E", c("1", "3", "9", "10", "14", "76")), function(x) 
+  max(plasmid.table[plasmid.table$PTU==x, "Size"])))
+```

### 2025-01-27 scripts/kmer_functions.py
scripts to oriengt leading region

@@ -1,0 +1,99 @@
+import argparse
+import itertools as iter  
+
+# For reverse complement
+alt_map = {'ins':'0'}
+complement = {'A': 'T', 'C': 'G', 'G': 'C', 'T': 'A'}
+
+def get_options():
+    parser = argparse.ArgumentParser(description='Counts k-mers in a fasta and returns in lexicographic order.',
+                                     prog='countKmers')
+    parser.add_argument('--fasta', help='fasta file', required=True)
+    parser.add_argument('--k', help='value of k', required=True) 
+    return parser.parse_args()
+
+
+def reverse_complement(seq):
+    for k,v in alt_map.items():
+        seq = seq.replace(k,v)
+    bases = list(seq)
+    bases = reversed([complement.get(base,base) for base in bases])
+    bases = ''.join(bases)
+    for k,v in alt_map.items():
+        bases = bases.replace(v,k)
+    return bases
+
+def all_palindromes(N):
+    '''Returns all palindromes of length N.'''
+    palindromes = []
+    for seq in iter.product('ATCG', repeat=N):
+        seq = ''.join(seq)
+        half_seq = seq[0:int(N/2)]
+        end_seq = seq[int(N/2):int(N)]
+        if half_seq==reverse_complement(end_seq):
+            palindromes.append(seq)
+    return palindromes
+
+
+def get_seq(input_fasta):
+    dna = ''
+    with open(input_fasta, 'r') as f:
+        for line in f.readlines():
+            if line.startswith('>'):
+                dna += 'Z'
+            else:
+                dna += line.strip('\n')
+    return(dna)
+
+# Using Matthew Fahrbach solution from Rosalind problem 
+# K-mer counts are enumerated in lexicographical order
+# of the k-mers
+# e.g. AAAA comes first for k=4
+def count_kmers(fasta, k, with_rev_comp=False):
+    dna = get_seq(fasta)
+    if with_rev_comp:
+        rev_comp_dna = reverse_complement(dna)
+        dna = dna+"Z"+rev_comp_dna
+    table = {};
+    for kmer in iter.product('ACGT', repeat=k):
+        table[''.join(kmer)] = 0
+
+    for i in range(len(dna) - k + 1):
+        kmer = dna[i:i + k]
+        if all([base in ['A', 'T', 'C', 'G'] for base in kmer]): # only if kmer is ATCG-based, no other characters allowed
[…]

### 2025-01-27 scripts/orient_leading_region.py

@@ -1,0 +1,217 @@
+# Orient plasmid by leading region
+
+# Libraries
+import argparse  
+from itertools import product
+import glob
+import pandas as pd
+import re
+from math import floor 
+import kmer_functions as kf # Useful k-mer functions 
+
+
+def get_options():
+    parser = argparse.ArgumentParser(description='Orient plasmid by leading region i.e. write fastas so that the first base of the sequence is the start of the leading region. Assumes circular plasmid!')
+    parser.add_argument('--output', help='output file', type=str)
+    parser.add_argument('--fasta_file_search', help='string to find files of plasmids with glob (can contain wildcards e.g. 06_plaspan/*/_fasta/*)', type=str)
+    parser.add_argument('--leading_region_file', help='file with locations of leading region', type=str)
+    parser.add_argument('--leading_region_size', help='leading region size (bp)', default=5000, required=False)
+    parser.add_argument('--sliding_window', help='sliding window size', default=False, required=False)
+    parser.add_argument('--step_size', help='step size for sliding window', default=False, required=False)
+    parser.add_argument('--k', help='palindrome length', default=False, required=False)
+    return parser.parse_args()
+
+
+alt_map = {'ins':'0'}
+complement = {'A': 'T', 'C': 'G', 'G': 'C', 'T': 'A'}
+
+def reverse_complement(seq):
+    for k,v in alt_map.items():
+        seq = seq.replace(k,v)
+    bases = list(seq)
+    bases = reversed([complement.get(base,base) for base in bases])
+    bases = ''.join(bases)
+    for k,v in alt_map.items():
+        bases = bases.replace(v,k)
+    return bases
+
+def get_plasmid_seq(input_fasta):
+    dna = ''
+    with open(input_fasta, 'r') as f:
+        for line in f.readlines():
+            if not line.startswith('>'):
+                dna += line.strip('\n')
+            else:
+                pass
+    return ''.join(char if char in 'ATGC' else 'N' for char in dna)
+
+def orientate_by_leading_region(sequence, start_position=1, orientation="downstream"):
+	return_sequence = ''
+	if orientation=="downstream":
+		return_sequence += sequence[(start_position-1):]
+		return_sequence +=  sequence[0:(start_position-1)]
[…]

### 2025-01-27 scripts/Paper-analysis.Rmd
adding in leading region analysis

@@
 library(xtable)
+library(dplyr)
 ```
@@
 # Minimum sizes
-sort(sapply(lr.identified.PTU.names, function(x) 
+PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
   min(plasmid.table[plasmid.table$PTU==x, "Size"])))
@@
 ```
+
+These PTUs are retained for analysing the leading region:
+
+```{r}
+cat(sort(names(PTU.min.sizes[PTU.min.sizes>20000])), sep="\n")
+```
+### Sliding window analysis
+
+```{r}
+# Theoretical expectation for 6-bp palindrome density
+theoreticalExpectation = function(p){# p is GC content
+  return ((p**2 - p + 1/2)**3)
+}
+d = read.csv('../test_output.csv', header=F, stringsAsFactors = F)
+d.reduced = d[which((d$V3 %% 500)==0), ]
+
+colnames(d.reduced) = c("PTU", "plasmid", "x", "density", "GC_n", "expected_density")
+d.reduced$expected_density_new_calc = theoreticalExpectation(d.reduced$GC_n/500)
+
+d.reduced$diff_density = d.reduced$density-d.reduced$expected_density_new_calc
+
+medians = d.reduced %>% group_by(PTU, plasmid) %>%
+  summarise(median=median(density)) %>% group_by(PTU) %>%
+  summarise(median=median(median))
+
+# One line per plasmid
+d.reduced.median = d.reduced %>% group_by(PTU, x) %>% 
+  summarise(median=median(diff_density),
+            n=length(diff_density), 
+            median.original=median(density),
+            GC=median(GC_n))
+
+# Maximum counts
+d.reduced.median.plot = d.reduced.median %>% 
+  group_by(PTU) %>%
+  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
+         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
+  ungroup() %>%  # Remove the grouping
+  select(-max_n)  # Optionally remove the max_n column if not needed
+# Plot the distribution
+p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("6-mer palindrome density (5kb window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+```

### 2025-01-27 scripts/orient_leading_region.py
remove old palindrome density funcs

@@ -75,21 +75,6 @@ def find_dna_palindromes(sequence, k):
             locations.append(i)
     return locations
 
-def palindrome_probability(k, gc_content):
-    # This is for probability of a random sequence being a palindrome - not the same as palindrome density in a random sequence
-    if k % 2 != 0:
-        # Odd-length palindromes are impossible in DNA because of base pairing
-        return 0
-    # Probability of a GC or CG pair
-    gc_prob = (gc_content / 2) ** 2
-    # Probability of an AT or TA pair
-    at_prob = ((1 - gc_content) / 2) ** 2
-    # For a palindrome, the first half of the sequence determines the second half
-    half_k = k // 2
-    # Probability that the first half forms a palindrome
-    palindrome_prob = (gc_prob + at_prob) ** half_k
-    return palindrome_prob
-
 
 def palindrome_density(sequence, k, window_size, step_size):
     palindrome_locations = find_dna_palindromes(sequence, k)
@@ -98,7 +83,9 @@ def palindrome_density(sequence, k, window_size, step_size):
         end = start + window_size
         count = sum(1 for loc in palindrome_locations if start <= loc < end)
         density = count / window_size
-        densities.append((start, density, sequence[start:end].count('G')+sequence[start:end].count('C'), palindrome_probability(6, float(sequence[start:end].count('G')+sequence[start:end].count('C'))/float(window_size))))
+        densities.append((start, 
+        	density, 
+        	sequence[start:end].count('G')+sequence[start:end].count('C')))
     return densities
 
 def leading_region(plasmid_fasta, 

### 2025-01-27 scripts/orient_leading_region.py
what script does

@@ -1,4 +1,4 @@
-# Orient plasmid by leading region
+# Orient plasmid by leading region and use sliding window for palindrome density
 
 # Libraries
 import argparse  

### 2025-01-27 scripts/Paper-analysis.Rmd
save k-mer density expectation figure

--- 2025-01-27 scripts/Paper-analysis.Rmd
@@
 ```{r}
-lr = read.csv('data/leading_region_081123.tsv', sep='\t')
-plasmid.table = read.csv('data/plasmid_table.tsv', sep='\t')
-retained.PTUs.lr = read.csv('data/leading-region-PTUs.txt', header=F)$V1
+lr = read.csv('../data/leading_region_081123.tsv', sep='\t')
+plasmid.table = read.csv('../data/plasmid_table.tsv', sep='\t')
+retained.PTUs.lr = read.csv('../data/leading-region-PTUs.txt', header=F)$V1
 lr.identified.PTU.names = sort(unique(lr$PTU))
 # Minimum sizes
-sort(sapply(lr.identified.PTU.names, function(x) 
+PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
   min(plasmid.table[plasmid.table$PTU==x, "Size"])))
@@
 ```
+
+These PTUs are retained for analysing the leading region:
+
+```{r}
+leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
+cat(leading.region.PTUs, sep="\n")
+```
+### Sliding window analysis
+
+k=6 
+```{r}
+# Theoretical expectation for 6-bp palindrome density
+theoreticalExpectation = function(p){# p is GC content
+  return ((p**2 - p + 1/2)**3)
+}
+d = read.csv('../test_output.csv', header=F, stringsAsFactors = F)
+colnames(d) = c("PTU", "plasmid", "x", "density", "GC_n")
+
+d.reduced = d[which(d$PTU %in% leading.region.PTUs), ]
+
+d.reduced$expected_density_new_calc = theoreticalExpectation(d.reduced$GC_n/5000)
+
+d.reduced$diff_density = d.reduced$density-d.reduced$expected_density_new_calc
+
+medians = d.reduced %>% group_by(PTU, plasmid) %>%
+  summarise(median=median(density)) %>% group_by(PTU) %>%
+  summarise(median=median(median))
+
+# One line per plasmid
+d.reduced.median = d.reduced %>% group_by(PTU, x) %>% 
+  summarise(median=median(diff_density),
+            n=length(diff_density), 
+            median.original=median(density),
+            GC=median(GC_n))
+
+# Maximum counts
+d.reduced.median.plot = d.reduced.median %>% 
+  group_by(PTU) %>%
+  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
+         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
+  ungroup() %>%  # Remove the grouping
+  select(-max_n)  # Optionally remove the max_n column if not needed
+# Plot the distribution
+p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("Deviation from expected 6-mer palindrome density (5kb sliding window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf'.
+       width=10, height=8)
+```
+
+Repeat for k=4?

### 2025-01-27 scripts/orient_leading_region.py
store number of palindromes too

@@ -78,15 +78,17 @@ def find_dna_palindromes(sequence, k):
 
 def palindrome_density(sequence, k, window_size, step_size):
     palindrome_locations = find_dna_palindromes(sequence, k)
-    densities = []
+    results = []
     for start in range(0, len(sequence) - window_size + 1, step_size):
         end = start + window_size
         count = sum(1 for loc in palindrome_locations if start <= loc < end)
         density = count / window_size
-        densities.append((start, 
+        results.append((start,
+        	window_size, 
+        	count,
         	density, 
         	sequence[start:end].count('G')+sequence[start:end].count('C')))
-    return densities
+    return results
 
 def leading_region(plasmid_fasta, 
 								start_position,

### 2025-01-27 scripts/Paper-analysis.Rmd
messing around with palindrome density expectation

--- 2025-01-27 scripts/Paper-analysis.Rmd
@@
 ```
+
+These PTUs are retained for analysing the leading region:
+
+```{r}
+leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
+cat(leading.region.PTUs, sep="\n")
+```
+### Sliding window analysis
+
+k=6 
+```{r}
+# Theoretical expectation for 6-bp palindrome density
+theoreticalPalindromeExpectation = function(p){# p is GC content
+  return ((p**2 - p + 1/2)**3)
+}
+d = read.csv('../test_output_window1000_step1000.csv', header=F, stringsAsFactors = F)
+colnames(d) = c("PTU", "plasmid", "x", "window_size", "palindrome_count", "palindrome_density", "GC_n")
+
+d.reduced = d[which(d$PTU %in% leading.region.PTUs), ]
+
+d.reduced$expected_density = theoreticalPalindromeExpectation(d.reduced$GC_n/d.reduced$window_size)
+
+d.reduced$diff_density = d.reduced$palindrome_density-d.reduced$expected_density
+
+# One line per plasmid
+d.reduced.median = d.reduced %>% group_by(PTU, x) %>% 
+  summarise(median=median(diff_density),
+            n=length(diff_density), 
+            median.original=median(palindrome_density),
+            GC=median(GC_n))
+
+# Maximum counts
+d.reduced.median.plot = d.reduced.median %>% 
+  group_by(PTU) %>%
+  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
+         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
+  ungroup() %>%  # Remove the grouping
+  select(-max_n)  # Optionally remove the max_n column if not needed
+# Plot the distribution
+p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("Deviation from expected 6-mer palindrome density (5kb sliding window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf',
+       width=10, height=8)
+```
+
+Repeat for k=4?

### 2025-01-27 scripts/Paper-analysis.Rmd
use latest choice of sliding windo / step size

@@
 }
-d = read.csv('../test_output_window1000_step1000.csv', header=F, stringsAsFactors = F)
+d = read.csv('../test_output_window5000_step500.csv', header=F, stringsAsFactors = F)
 colnames(d) = c("PTU", "plasmid", "x", "window_size", "palindrome_count", "palindrome_density", "GC_n")
@@
 ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf',
-       width=10, height=8)
+       width=10, height=6)
 ```

### 2025-01-28 scripts/Paper-analysis.Rmd
sliding window figures

@@
 # Plot the distribution
-p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median))+
+p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median*1000))+
   geom_hline(yintercept = 0, linetype='dashed')+
@@
   facet_wrap(~PTU, scales="free_x", nrow=3)+
-  ylab("Deviation from expected 6-mer palindrome density (5kb sliding window)")+
+  ylab("Palindrome depletion per kb (5kb sliding window)")+
   labs(alpha="% of plasmids within PTU")+
@@
        width=10, height=6)
+# Plot only first 25kb
+ggsave(p.lines.expectation+scale_x_continuous(labels = function(x) x / 1000, limits=c(0,50000)), file='../results/6mer-deviation-expectation-0-50-kb.pdf',
+       width=10, height=6)
+
 ```
 
-Repeat for k=4?
+### Compare leading region to rest of plasmid
+
+```{r}
+# Start from d.reduced, which contains per-plasmid results with 5000bp sliding window and step size of 500bp
+# Compare depletion in first 5kb with average for the rest of the plasmid
+diff.leading.region.average = d.reduced %>% 
+  group_by(plasmid, PTU) %>%
+  mutate(diff.count=palindrome_count[x=="0"]-mean(palindrome_count),
+         diff.density=diff_density[x=="0"]-mean(diff_density)) %>%
+  filter(x=="0")
+
+# Order the PTUs by mean values of depletion for better visual display
+mean.values = diff.leading.region.average %>% group_by(PTU) %>%
+  summarise(mean=mean(diff.density)*1000)
+diff.leading.region.average$PTU = ordered(diff.leading.region.average$PTU,
+                                          levels=mean.values$PTU[order(mean.values$mean)])
+p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
+  ggbeeswarm::geom_quasirandom(alpha=0.8)+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  geom_hline(yintercept = 0)+
+  xlab("")+
+  theme(axis.text=element_text(colour="black"))+
+  ylab("Palindrome depletion in leading region (per kb)")+
+  coord_flip()
+p.leading.region.diff.density
+# Based on this, remove two outliers
+p.leading.region.diff.density = p.leading.region.diff.density+
+  ylim(c(-10,5))
+ggsave(p.leading.region.diff.density, file='../results/6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
+       width=5, height=4)
+```

### 2025-01-28 scripts/count_kmers_leading_region.py
script to count given k-mer targets across plasmid

@@ -1,0 +1,128 @@
+# Count abundance of targets in leading region of plasmid vs. rest of plasmid
+
+# Libraries
+import argparse  
+import itertools as iter
+import glob
+import pandas as pd
+import re
+from math import floor 
+import kmer_functions as kf # Useful k-mer functions 
+
+
+def get_options():
+    parser = argparse.ArgumentParser(description='Orient plasmid by leading region i.e. write fastas so that the first base of the sequence is the start of the leading region. Assumes circular plasmid!')
+    parser.add_argument('--output', help='output file', type=str)
+    parser.add_argument('--fasta_file_search', help='string to find files of plasmids with glob (can contain wildcards e.g. 06_plaspan/*/_fasta/*)', type=str)
+    parser.add_argument('--leading_region_file', help='file with locations of leading region', type=str)
+    parser.add_argument('--leading_region_size', help='leading region size (bp)', default=5000, required=False)
+    parser.add_argument('--sliding_window', help='sliding window size', default=False, required=False)
+    parser.add_argument('--step_size', help='step size for sliding window', default=False, required=False)
+    parser.add_argument('--k', help='palindrome length', default=False, required=False)
+    return parser.parse_args()
+
+
+alt_map = {'ins':'0'}
+complement = {'A': 'T', 'C': 'G', 'G': 'C', 'T': 'A'}
+
+def reverse_complement(seq):
+    for k,v in alt_map.items():
+        seq = seq.replace(k,v)
+    bases = list(seq)
+    bases = reversed([complement.get(base,base) for base in bases])
+    bases = ''.join(bases)
+    for k,v in alt_map.items():
+        bases = bases.replace(v,k)
+    return bases
+
+def get_plasmid_seq(input_fasta):
+    dna = ''
+    with open(input_fasta, 'r') as f:
+        for line in f.readlines():
+            if not line.startswith('>'):
+                dna += line.strip('\n')
+            else:
+                pass
+    return ''.join(char if char in 'ATGC' else 'N' for char in dna)
+
+def orientate_by_leading_region(sequence, start_position=1, orientation="downstream"):
+	return_sequence = ''
+	if orientation=="downstream":
+		return_sequence += sequence[(start_position-1):]
[…]

### 2025-01-28 scripts/split_leading_and_lagging.py
script to produce leading/lagging regions from plasmid sequence

@@ -1,0 +1,85 @@
+
+import argparse
+import glob
+import pandas as pd
+import re
+
+def get_options():
+    parser = argparse.ArgumentParser(description='Orient plasmid by leading region then split into first and last X kb (assumes circular plasmid!')    
+    parser.add_argument('--output_dir', help='output directory', type=str)
+    parser.add_argument('--fasta_file_search', help='string to find files of plasmids with glob (can contain wildcards e.g. 06_plaspan/*/_fasta/*)', type=str)
+    parser.add_argument('--leading_region_file', help='file with locations of leading region', type=str)
+    parser.add_argument('--leading_region_size', help='leading region size (bp)', default=5000, required=False)
+    return parser.parse_args()
+
+alt_map = {'ins':'0'}
+complement = {'A': 'T', 'C': 'G', 'G': 'C', 'T': 'A'}
+
+def reverse_complement(seq):
+    for k,v in alt_map.items():
+        seq = seq.replace(k,v)
+    bases = list(seq)
+    bases = reversed([complement.get(base,base) for base in bases])
+    bases = ''.join(bases)
+    for k,v in alt_map.items():
+        bases = bases.replace(v,k)
+    return bases
+
+def get_plasmid_seq(input_fasta):
+    dna = ''
+    with open(input_fasta, 'r') as f:
+        for line in f.readlines():
+            if not line.startswith('>'):
+                dna += line.strip('\n')
+            else:
+                pass
+    return ''.join(char if char in 'ATGC' else 'N' for char in dna)
+
+def orientate_by_leading_region(sequence, start_position=1, orientation="downstream"):
+	return_sequence = ''
+	if orientation=="downstream":
+		return_sequence += sequence[(start_position-1):]
+		return_sequence +=  sequence[0:(start_position-1)]
+	elif orientation=="upstream":
+		return_sequence += reverse_complement(sequence[0:(start_position-1)])
+		return_sequence += reverse_complement(sequence[(start_position-1):])
+	return return_sequence
+
+def seq_with_leading_region_at_start(seq, 
+								start_position,
+								orientation):
+	plasmid_seq = seq
+	new_seq = orientate_by_leading_region(plasmid_seq, 
+		int(start_position), orientation)
+	return(new_seq)
+
+
[…]

### 2025-01-28 scripts/split_leading_and_lagging.py
ignore double entires

@@ -71,15 +71,16 @@ if __name__=="__main__":
 		plasmid_name = re.sub(".fasta", "", re.sub(".*\\/", "", fasta_file))
 		ptu_name = re.sub(".*06_plaspan\\/", "", re.sub("\\/_fasta.*", "", fasta_file))
 		if plasmid_name in leading_region_df.index:
-			print(plasmid_name, ptu_name)
-			leading_region_string = leading_region_df.loc[plasmid_name]["Leading Region"]
-			leading_region_start = re.sub(" in.*", "", re.sub("From ", "", leading_region_string))
-			#print(leading_region_start)
-			leading_region_orientation = re.sub(".*in ", "", re.sub(" direction.", "", leading_region_string))
-			new_seqs = leading_and_lagging(fasta_file, leading_region_start, leading_region_orientation, size=size_region)
-			with open(output_dir+'/'+plasmid_name+'_leading_'+str(size_region)+'.fa', 'w') as f:
-				f.write('>test\n%s' % new_seqs[0])
-			with open(output_dir+'/'+plasmid_name+'_lagging_'+str(size_region)+'.fa', 'w') as f:
-				f.write('>test\n%s' % new_seqs[1])
+			if len(leading_region_df.loc[plasmid_name])!=2: # ignore double oriT
+print(plasmid_name, ptu_name)
+				leading_region_string = leading_region_df.loc[plasmid_name]["Leading Region"]
+				leading_region_start = re.sub(" in.*", "", re.sub("From ", "", leading_region_string))
+				#print(leading_region_start)
+				leading_region_orientation = re.sub(".*in ", "", re.sub(" direction.", "", leading_region_string))
+				new_seqs = leading_and_lagging(fasta_file, leading_region_start, leading_region_orientation, size=size_region)
+				with open(output_dir+'/'+plasmid_name+'_leading_'+str(size_region)+'.fa', 'w') as f:
+					f.write('>test\n%s' % new_seqs[0])
+				with open(output_dir+'/'+plasmid_name+'_lagging_'+str(size_region)+'.fa', 'w') as f:
+					f.write('>test\n%s' % new_seqs[1])
 
 

### 2025-01-28 scripts/split_leading_and_lagging.py
indentation fix

@@ -72,7 +72,7 @@ if __name__=="__main__":
 		ptu_name = re.sub(".*06_plaspan\\/", "", re.sub("\\/_fasta.*", "", fasta_file))
 		if plasmid_name in leading_region_df.index:
 			if len(leading_region_df.loc[plasmid_name])!=2: # ignore double oriT
-print(plasmid_name, ptu_name)
+				print(plasmid_name, ptu_name)
 				leading_region_string = leading_region_df.loc[plasmid_name]["Leading Region"]
 				leading_region_start = re.sub(" in.*", "", re.sub("From ", "", leading_region_string))
 				#print(leading_region_start)

### 2025-01-28 scripts/rmes-score-discrepancy.py
rename core/accessory to file1/file2 for generality

@@ -5,13 +5,13 @@ import argparse
 
 def read_args():
     parser = argparse.ArgumentParser(description="Calculate median discrepancies and perform Wilcoxon test.")
-    parser.add_argument("--core", required=True, help="Path to the core CSV file.")
-    parser.add_argument("--accessory", required=True, help="Path to the accessory CSV file.")
+    parser.add_argument("--file1", required=True, help="Path to the first CSV file (e.g. core genes).")
+    parser.add_argument("--file2", required=True, help="Path to the other CSV file (e.g. accessory genes).")
     parser.add_argument("--kmers", required=True, help="Path to the kmers text file.")
 
     return parser.parse_args()
 
-# Load data from core and accessory files
+# Load data from a csv file
 def load_csv(file):
     return pd.read_csv(file)
 
@@ -20,19 +20,19 @@ def load_kmers(file):
     with open(file, 'r') as f:
         return set(line.strip().lower() for line in f)
 
-def calculate_discrepancy_and_test(core_file, accessory_file, kmers_file):
-    # Load core and accessory data
-    core = load_csv(core_file)
-    accessory = load_csv(accessory_file)
+def calculate_discrepancy_and_test(first_file, second_file, kmers_file):
+    # Load the data from both files
+    df1 = load_csv(first_file)
+    df2 = load_csv(second_file)
 
     # Load k-mers
     kmers = load_kmers(kmers_file)
 
-    # Merge core and accessory on "word"
-    merged = pd.merge(core, accessory, on='word', suffixes=('_core', '_accessory'))
+    # Merge on "word"
+    merged = pd.merge(df1, df2, on='word', suffixes=('_file1', '_file2'))
 
     # Calculate score discrepancy
-    merged['score_discrepancy'] = merged['score_core'] - merged['score_accessory']
+    merged['score_discrepancy'] = merged['score_file1'] - merged['score_file2']
 
     # Separate into k-mers and other words
     kmers_data = merged[merged['word'].isin(kmers)]
@@ -59,8 +59,8 @@ def calculate_discrepancy_and_test(core_file, accessory_file, kmers_file):
 # Example usage
 if __name__ == "__main__":
     args = read_args()
-    results = calculate_discrepancy_and_test(args.core, args.accessory, args.kmers)
-    filename_base = args.core.split('_')[0].rsplit("/", 1)[-1]
+    results = calculate_discrepancy_and_test(args.file1, args.file2, args.kmers)
[…]

### 2025-01-28 scripts/rmes-score-discrepancy.py
note to make name extraction more general

@@ -60,7 +60,7 @@ def calculate_discrepancy_and_test(first_file, second_file, kmers_file):
 if __name__ == "__main__":
     args = read_args()
     results = calculate_discrepancy_and_test(args.file1, args.file2, args.kmers)
-    filename_base = args.file1.split('_')[0].rsplit("/", 1)[-1]
+    filename_base = args.file1.split('_')[0].rsplit("/", 1)[-1] # Need to replace this with more general approach
 
     results = [filename_base,results['n_special_kmers'], results['median_discrepancy_kmers'], results['median_discrepancy_others'], results['wilcoxon_statistic'],results['p_value']]
     print(','.join([str(x) for x in results]))

### 2025-01-29 scripts/Paper-analysis.Rmd
more analysis

@@
   summarise(mean=mean(diff.density)*1000)
-diff.leading.region.average$PTU = ordered(diff.leading.region.average$PTU,
-                                          levels=mean.values$PTU[order(mean.values$mean)])
+diff.leading.region.average$PTU = ordered(gsub("PTU-", "", diff.leading.region.average$PTU),
+                                          levels=gsub("PTU-", "", mean.values$PTU[order(mean.values$mean)]))
 p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
@@
   xlab("")+
-  theme(axis.text=element_text(colour="black"))+
+  theme(axis.text=element_text(colour="black", size=18))+
   ylab("Palindrome depletion in leading region (per kb)")+
-  coord_flip()
+  theme(axis.text.x=element_text(angle=45, hjust=1))+
+  ylab("")
 p.leading.region.diff.density
@@
 p.leading.region.diff.density = p.leading.region.diff.density+
-  ylim(c(-10,5))
+  scale_y_continuous(breaks=seq(-8, 4, 2), limits=c(-8,5))
 ggsave(p.leading.region.diff.density, file='../results/6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
-       width=5, height=4)
+       width=5.5, height=6)
 ```
+### Averaging over all plasmids
+
+Similar to Samuel et al.'s display of anti-defence gene density, average over all conjugative plasmids.
+
+```{r}
+overall.median = d.reduced %>% filter(x<50000) %>%
+  group_by(x) %>%
+  summarise(y=median(diff_density),
+            n=length(diff_density))
+p.overall.average = ggplot(overall.median, aes(x, y*1000))+
+  geom_line()+
+  scale_y_continuous(breaks=seq(-10, 0, 2), limits = c(-9, 0))+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  ylab("Palindrome depletion (per kb)")+
+    scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(axis.text=element_text(colour="black"))
+p.overall.average
+# Save
+ggsave(p.overall.average, file='../results/6mer-deviation-overall.pdf', 
+       width=3, height=3)
+```
+
+```{r}
+lagging.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_lagging_10000_k6.csv', header=T)
+leading.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_leading_10000_k6.csv', header=T)
+merged.df = merge(leading.df, lagging.df, by="word")
+palindromes = tolower(read.csv('~/Downloads/trieste/k6.txt', header=F)$V1)
+ggplot(merged.df[which(merged.df$word %in% palindromes),], aes(score.x, score.y))+
+  geom_point()
+merged.df.palindromes = merged.df[which(merged.df$word %in% palindromes),]
+table(merged.df.palindromes$score.x<merged.df.palindromes$score.y)
+kmer_targets = tolower(read.csv('~/Downloads/trieste/PTU-LM_targets.txt', header=F)$V1)
+ggplot(merged.df[which(merged.df$word %in% kmer_targets),], aes(score.x, score.y))+
+  geom_point()
+merged.df.targets = merged.df[which(merged.df$word %in% kmer_targets),]
+table(merged.df.targets$score.x<merged.df.targets$score.y)
+
+```

### 2025-01-29 scripts/Paper-analysis.Rmd
leading vs. lagging palindrome scores

@@
 ```
+
+### Palindromes in leading vs. lagging
+
+```{r}
+scores_df = read.csv('../test_6mer_leading_lagging_scores.csv', header=T)
+scores_average <- scores_df %>%
+  rowwise() %>%
+  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
+  ungroup() %>%
+  select(kmer, score_avg)  
+# Order by average score
+scores_average$kmer = ordered(scores_average$kmer,
+                              levels=scores_average$kmer[order(scores_average$score_avg)])
+scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
+  coord_flip()
+
+# Same for counts
+counts_df = read.csv('../test_6mer_leading_lagging_counts.csv', header=T)
+counts_average <- counts_df %>%
+  rowwise() %>%
+  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
+  ungroup() %>%
+  select(kmer, count_avg)
+
+# Merge together
+merge_scores_counts = merge(counts_average, scores_average, by="kmer")
+# average score and average count are well-correlated
+ggplot(merge_scores_counts, aes(count_avg, score_avg))+geom_point()
+
+# Order by average score
+merge_scores_counts$kmer = ordered(merge_scores_counts$kmer,
+                              levels=scores_average$kmer[order(scores_average$score_avg)])
+scores.plot = ggplot(merge_scores_counts, aes(score_avg, kmer))+
+    geom_vline(xintercept = 0)+
+  geom_bar(stat="identity")+
+  geom_vline(xintercept = median(merge_scores_counts$score_avg), linetype='dashed', colour='red')+
+  theme_bw()+
+  theme(axis.text=element_text(colour="black", size=12),
+        axis.title.x=element_text(size=24),
+        axis.text.x=element_text(size=18))+
+  theme(axis.text.y=element_text(hjust=0))+
+  ylab("")+
+  xlab("Mean depletion score\n(leading vs. lagging, 10kb)")
+wilcox.test(merge_scores_counts$score_avg)
+ggsave(scores.plot, file='../results/6mer-palindromes-scores.pdf',
+       height=10, width=10)
+```

### 2025-01-29 scripts/rmes-score-discrepancy-leading-lagging.py

@@ -1,0 +1,85 @@
+import pandas as pd
+import scipy.stats as stats
+import numpy as np
+import argparse 
+import re
+import glob
+import os
+
+def read_args():
+    parser = argparse.ArgumentParser(description="Merge results for .")
+    parser.add_argument("--directory", required=True, help="Search path for output file.")
+    parser.add_argument("--kmers", required=True, help="Path to the kmers text file.")
+    parser.add_argument("--k", required=True, help="Value of k.")
+    parser.add_argument("--output", required=True, help="Output file prefix.")
+
+    return parser.parse_args()
+
+# Load data from a csv file
+def load_csv(file):
+    return pd.read_csv(file)
+
+# Load kmers from kmers.txt
+def load_kmers(file):
+    with open(file, 'r') as f:
+        return set(line.strip().lower() for line in f)
+
+def merge_score_files(first_file, second_file, kmer_file):
+    # Load the data from both files
+    df1 = load_csv(first_file)
+    df2 = load_csv(second_file)
+    kmers = load_kmers(kmer_file)
+
+    # Merge on "word"
+    merged = pd.merge(df1, df2, on='word', suffixes=('_file1', '_file2'))
+
+    # Calculate score discrepancy
+    merged['score_discrepancy'] = merged['score_file1'] - merged['score_file2']
+    merged['count_discrepancy'] = merged['count_file1'] - merged['count_file2']
+
+    merged = merged[merged['word'].isin(kmers)]
+    merged = merged[['word', 'score_discrepancy', 'count_discrepancy']]
+    sorted_df = merged.sort_values(by='word')
+    return [list(sorted_df['score_discrepancy']) , list(sorted_df['count_discrepancy'])]
+
+# For each plasmid, compute the discrepancies and then return the results 
+# save them in a dictionary with motifs as keys
+def merge_comparisons(directory, kmers_file, k):
+    # Find all leading and lagging files in the given directory
+    leading_files = sorted(glob.glob(os.path.join(directory, "*leading*"+str(k)+".csv")))
+    lagging_files = sorted(glob.glob(os.path.join(directory, "*lagging*"+str(k)+".csv")))
+    kmers = load_kmers(kmers_file)
+    all_scores = []
+    all_counts = []
+    for leading_file, lagging_file in zip(leading_files, lagging_files):
+        print(leading_file, lagging_file)
+        results = merge_score_files(leading_file, lagging_file, kmers_file)
[…]

### 2025-01-29 scripts/Paper-analysis.Rmd
more notes on analysis

@@
 cat(leading.region.PTUs, sep="\n")
+```
+
+For convenience for other scripts, I save some files with names of plasmids within each PTU.
+
+```{r}
+for (PTU in unique(plasmid.table$PTU)){
+  cat(plasmid.table$ID[plasmid.table$PTU==PTU], sep="\n",
+      file=paste0('~/Downloads/trieste/', PTU, '_plasmids.txt'))
+}
+
 ```
@@
 ```
+
+Test for a particular PTU.
+
+```{r}
+
+scores_df = read.csv('../test_bokz_k5_scores.csv', header=T)
+scores_average <- scores_df %>%
+  rowwise() %>%
+  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
+  ungroup() %>%
+  select(kmer, score_avg)  
+# Order by average score
+scores_average$kmer = ordered(scores_average$kmer,
+                              levels=scores_average$kmer[order(scores_average$score_avg)])
+scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
+  coord_flip()
+```

### 2025-01-29 scripts/rmes-score-discrepancy-leading-lagging.py

@@ -38,6 +38,7 @@ def merge_score_files(first_file, second_file, kmer_file):
     merged['count_discrepancy'] = merged['count_file1'] - merged['count_file2']
 
     merged = merged[merged['word'].isin(kmers)]
+    kmers_used = merged['word']
     merged = merged[['word', 'score_discrepancy', 'count_discrepancy']]
     sorted_df = merged.sort_values(by='word')
     return [list(sorted_df['score_discrepancy']) , list(sorted_df['count_discrepancy'])]
@@ -66,19 +67,22 @@ if __name__ == "__main__":
     kmers = sorted(load_kmers(args.kmers))
     print(kmers)
     results = merge_comparisons(args.directory, args.kmers, args.k)
+    # Only use k-mers of right length
+    kmers_reduced = [k for k in kmers if len(k)==int(args.k)]
+
     scores = results[0]
     counts = results[1]
     # Convert scores list of lists into a DataFrame
     scores_df = pd.DataFrame(scores).T  # Transpose to align scores with kmers
     # Assign column names
     scores_df.columns = [f'score{i+1}' for i in range(len(scores))]
-    scores_df.insert(0, 'kmer', kmers)  # Insert kmers as the first column
+    scores_df.insert(0, 'kmer', kmers_reduced)  # Insert kmers as the first column
     scores_df.to_csv(args.output+'_scores.csv', index=False)
 
     counts_df = pd.DataFrame(counts).T  # Transpose to align scores with kmers
     # Assign column names
     counts_df.columns = [f'count{i+1}' for i in range(len(scores))]
-    counts_df.insert(0, 'kmer', kmers)  # Insert kmers as the first column
+    counts_df.insert(0, 'kmer', kmers_reduced)  # Insert kmers as the first column
     counts_df.to_csv(args.output+'_counts.csv', index=False)
 
 

### 2025-01-29 scripts/split_leading_and_lagging.py

@@ -8,7 +8,7 @@ def get_options():
     parser = argparse.ArgumentParser(description='Orient plasmid by leading region then split into first and last X kb (assumes circular plasmid!')    
     parser.add_argument('--output_dir', help='output directory', type=str)
     parser.add_argument('--fasta_file_search', help='string to find files of plasmids with glob (can contain wildcards e.g. 06_plaspan/*/_fasta/*)', type=str)
-    parser.add_argument('--leading_region_file', help='file with locations of leading region', type=str)
+	parser.add_argument('--leading_region_file', help='file with locations of leading region', type=str)
     parser.add_argument('--leading_region_size', help='leading region size (bp)', default=5000, required=False)
     return parser.parse_args()
 

### 2025-01-29 scripts/Paper-analysis.Rmd
new analysis leading vs. lagging

@@
 
-scores_df = read.csv('../test_bokz_k5_scores.csv', header=T)
+results = NULL
+for (PTU in leading.region.PTUs){
+  for (k in c(5,6)){
+    scores_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_scores.csv'), header=T)
+    n_targets = nrow(scores_df)
     scores_average <- scores_df %>%
@@
   select(kmer, score_avg)
-# Order by average score
-scores_average$kmer = ordered(scores_average$kmer,
-                              levels=scores_average$kmer[order(scores_average$score_avg)])
-scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
-  coord_flip()
+    results = rbind(results, c(PTU, k, n_targets, median(scores_average$score_avg)))
+  }
+}
+
+results = data.frame(results)
+colnames(results) = c("PTU", "k", "n.targets", "median.score")
+results$median.score= as.numeric(results$median.score)
+
+# Order by median score for k=6
+order.k.6.scores = results[which(results$k==6), "PTU"][order(results[which(results$k==6), "median.score"])]
+results$PTU = ordered(results$PTU,
+                      levels=rev(order.k.6.scores))
+p.depletion.targets = ggplot(results, aes(median.score, as.factor(PTU)))+
+  geom_vline(xintercept = 0, size=1)+
+  geom_bar(stat="identity", position="dodge")+
+  facet_wrap(~k, scales="free_x")+
+  ylab("")+
+  xlab("Median depletion score of within-range RM targets\n(leading vs. lagging, 10kb)")+
+   theme_bw()+
+  theme(panel.grid = element_blank())+
+  theme(axis.text=element_text(colour="black", size=12),
+        axis.title.x=element_text(size=24),
+        axis.text.x=element_text(size=18))+
+  theme(axis.text.y=element_text(hjust=0))
+ggsave(p.depletion.targets, file='../results/figure-depletion-within-range-targets.pdf',
+       width=10, height=6)
 ```

### 2025-01-29 scripts/run-rmes-score-discrepancy.py

@@ -1,52 +1,0 @@
-import subprocess
-import os
-import glob
-
-def run_comparisons(base_script, directory, kmers_file, output_file, k):
-    # Find all core and accessory files in the given directory
-    core_files = sorted(glob.glob(os.path.join(directory, "PTU-*_core_genes_k"+str(k)+".csv")))
-    accessory_files = sorted(glob.glob(os.path.join(directory, "PTU-*_accessory_genes_k"+str(k)+".csv")))
-    if kmers_file=="per-PTU":
-        kmers_files = sorted(glob.glob(os.path.join(directory, "PTU-*_targets.txt")))
-    # Ensure matching pairs
-    if len(core_files) != len(accessory_files):
-        print("Error: Mismatched number of core and accessory files.")
-        return
-
-    with open(output_file, "w") as out:
-        out.write("PTU,n_selected_kmers,discrepancy_selected,discrepancy_others,wilcoxon_stat,p_value\n")
-        for core_file, accessory_file, kmers_file in zip(core_files, accessory_files, kmers_files):
-                # Construct the command
-                command = [
-                    "python", base_script,
-                    "--file1", core_file,
-                    "--file2", accessory_file,
-                    "--kmers", kmers_file
-                ]
-
-                # Execute the command
-                result = subprocess.run(command, capture_output=True, text=True)
-
-                # Output the results
-                if result.returncode == 0:
-                    out.write(result.stdout)
-                else:
-                    print(f"Error for core: {core_file} and accessory: {accessory_file}")
-                    print(result.stderr)
-
-if __name__ == "__main__":
-    import argparse
-
-    # Argument parsing
-    parser = argparse.ArgumentParser(description="Run comparisons for multiple core and accessory files.")
-    parser.add_argument("--script", required=True, help="Path to the comparison script.")
-    parser.add_argument("--directory", required=True, help="Directory containing core and accessory files.")
-    parser.add_argument("--kmers", required=True, help="Path to the kmers text file (make sure it matches value of k). If 'per-PTU' is passed then will use targets by PTU, generated by scripts/create-PTU-target-DB.py")
-    parser.add_argument("--output", required=True, help="Path to the output file.")
-    parser.add_argument("--k", required=True, help="Value of k (4,5,6).")
-
-
-    args = parser.parse_args()
-
-    # Run comparisons
-    run_comparisons(args.script, args.directory, args.kmers, args.output, args.k)

### 2025-01-29 scripts/rmes-score-discrepancy-leading-lagging.py
delete print statements

@@ -53,7 +53,6 @@ def merge_comparisons(directory, kmers_file, k):
     all_scores = []
     all_counts = []
     for leading_file, lagging_file in zip(leading_files, lagging_files):
-        print(leading_file, lagging_file)
         results = merge_score_files(leading_file, lagging_file, kmers_file)
         all_scores.append(results[0])
         all_counts.append(results[1])
@@ -65,7 +64,6 @@ def merge_comparisons(directory, kmers_file, k):
 if __name__ == "__main__":
     args = read_args()
     kmers = sorted(load_kmers(args.kmers))
-    print(kmers)
     results = merge_comparisons(args.directory, args.kmers, args.k)
     # Only use k-mers of right length
     kmers_reduced = [k for k in kmers if len(k)==int(args.k)]

### 2025-01-29 scripts/Paper-analysis.Rmd
add counts

--- 2025-01-29 scripts/Paper-analysis.Rmd
@@
 library(dplyr)
+library(tidyr)
 ```
@@
   select(kmer, score_avg)
-# Order by average score
-scores_average$kmer = ordered(scores_average$kmer,
-                              levels=scores_average$kmer[order(scores_average$score_avg)])
-scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
-  coord_flip()
+    counts_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_counts.csv'), header=T)
+    counts_average <- counts_df %>%
+  rowwise() %>%
+  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
+  ungroup() %>%
+  select(kmer, count_avg)
+    results = rbind(results, c(PTU, k, n_targets, median(scores_average$score_avg), median(counts_average$count_avg)))
+  }
+}
+
+results = data.frame(results)
+colnames(results) = c("PTU", "k", "n.targets", "median.score", "median.count")
+results$median.score= as.numeric(results$median.score)
+results$median.count = as.numeric(results$median.count)
+
+# Order by median score for k=6
+order.k.6.scores = results[which(results$k==6), "PTU"][order(results[which(results$k==6), "median.score"])]
+results$PTU = ordered(results$PTU,
+                      levels=rev(order.k.6.scores))
+# Add number of targets to each
+
+p.depletion.targets = ggplot(results, aes(median.score, as.factor(PTU)))+
+  geom_vline(xintercept = 0, size=1)+
+  geom_bar(stat="identity", position="dodge")+
+  facet_wrap(~k, scales="free_x")+
+  ylab("")+
+  xlab("Median depletion score of within-range RM targets\n(leading vs. lagging, 10kb)")+
+   theme_bw()+
+  theme(panel.grid = element_blank())+
+  theme(axis.text=element_text(colour="black", size=12),
+        axis.title.x=element_text(size=24),
+        axis.text.x=element_text(size=18))+
+  theme(axis.text.y=element_text(hjust=0))
+ggsave(p.depletion.targets, file='../results/figure-depletion-within-range-targets.pdf',
+       width=10, height=6)
+
+# Is there a correlation between k=5 and k=6 results?
+results.df = results %>% select(k, median.score, PTU) %>%
+  pivot_wider(
+    names_from = k,  # Use values of 'k' as new column names
+    values_from = median.score  # Use 'median.score' for the column values
+)
 ```

### 2025-01-31 scripts/deambiguate-target-db.py
add main

@@ -44,19 +44,19 @@ def possibleSequences(dna_sequence):
             possible_strings = [x for addition in additions for x in addition]
     return possible_strings
 
+if __name__ == "__main__":
+    target_db = pd.read_csv("data/RM_target_db.csv")
+    new_target_db = []
+    for i, sequence in enumerate(target_db['sequence']):
+        new_sequences = possibleSequences(sequence)
+        row = list(target_db.loc[i])
+        for s in new_sequences:
+            new_target_db.append(row+[s])
 
-target_db = pd.read_csv("data/RM_target_db.csv")
-new_target_db = []
-for i, sequence in enumerate(target_db['sequence']):
-    new_sequences = possibleSequences(sequence)
-    row = list(target_db.loc[i])
-    for s in new_sequences:
-        new_target_db.append(row+[s])
-
-new_target_db_df = pd.DataFrame(new_target_db)
-new_target_db_df.columns = ['genus', 'sequence', 'genome', 'total_genomes', 'normalized_count', 'deambiguated_sequence']
-new_target_db_df = new_target_db_df.assign(length=[len(x) for x in new_target_db_df['deambiguated_sequence']])
-new_target_db_df.to_csv("data/RM_target_db_deambiguated.csv", index=False)
+    new_target_db_df = pd.DataFrame(new_target_db)
+    new_target_db_df.columns = ['genus', 'sequence', 'genome', 'total_genomes', 'normalized_count', 'deambiguated_sequence']
+    new_target_db_df = new_target_db_df.assign(length=[len(x) for x in new_target_db_df['deambiguated_sequence']])
+    new_target_db_df.to_csv("data/RM_target_db_deambiguated.csv", index=False)
 
 
 

### 2025-01-31 scripts/rmes-score-discrepancy-leading-lagging.py
work with either kmer file or do for all kmers

@@ -5,11 +5,12 @@ import argparse
 import re
 import glob
 import os
+import itertools as iter 
 
 def read_args():
     parser = argparse.ArgumentParser(description="Merge results for .")
     parser.add_argument("--directory", required=True, help="Search path for output file.")
-    parser.add_argument("--kmers", required=True, help="Path to the kmers text file.")
+    parser.add_argument("--kmers", required=False, help="Path to the kmers text file.", default="")
     parser.add_argument("--k", required=True, help="Value of k.")
     parser.add_argument("--output", required=True, help="Output file prefix.")
 
@@ -24,11 +25,14 @@ def load_kmers(file):
     with open(file, 'r') as f:
         return set(line.strip().lower() for line in f)
 
-def merge_score_files(first_file, second_file, kmer_file):
+def merge_score_files(first_file, second_file, kmer_file=False):
     # Load the data from both files
     df1 = load_csv(first_file)
     df2 = load_csv(second_file)
-    kmers = load_kmers(kmer_file)
+    if kmer_file != False:
+        kmers = load_kmers(kmer_file)
+    else:
+        kmers = sorted(df1['word']) # use all k-mers if no k-mer file
 
     # Merge on "word"
     merged = pd.merge(df1, df2, on='word', suffixes=('_file1', '_file2'))
@@ -45,42 +49,50 @@ def merge_score_files(first_file, second_file, kmer_file):
 
 # For each plasmid, compute the discrepancies and then return the results 
 # save them in a dictionary with motifs as keys
-def merge_comparisons(directory, kmers_file, k):
+def merge_comparisons(directory, k, kmers_file):
     # Find all leading and lagging files in the given directory
     leading_files = sorted(glob.glob(os.path.join(directory, "*leading*"+str(k)+".csv")))
     lagging_files = sorted(glob.glob(os.path.join(directory, "*lagging*"+str(k)+".csv")))
-    kmers = load_kmers(kmers_file)
+
     all_scores = []
     all_counts = []
     for leading_file, lagging_file in zip(leading_files, lagging_files):
         results = merge_score_files(leading_file, lagging_file, kmers_file)
         all_scores.append(results[0])
         all_counts.append(results[1])
-    return [all_scores, all_counts]
+    return [leading_files, all_scores, all_counts]
 
 
 
 # Example usage
 if __name__ == "__main__":
     args = read_args()
[…]

### 2025-01-31 scripts/deambiguate_rebase_seqs.py
deambiguate REBASE sequences

@@ -1,0 +1,26 @@
+# Deambiguate REBASE sequences
+# Created all_rebase_targets_250131 from:
+# 1. Downloading https://rebase.neb.com/rebase/link_orgref
+# 2. Extracted targets with grep "<5>""
+# But they need cleaning and deambiguating, which this script does
+import re 
+import deambiguate_target_db as de
+
+def clean_sequence(seq):
+    return re.sub(r'[\(\)\d/^,\s?]', '', seq)
+
+
+rebase_sequences = [clean_sequence(line.strip('\n')) for line in open('~/Downloads/trieste//all_rebase_targets_250131.txt')]
+
+unambiguous_sequences = []
+for s in rebase_sequences:
+	if len(s)<7:
+		new_seqs = de.possibleSequences(s)
+		for new_s in new_seqs:
+			unambiguous_sequences.append(new_s)
+
+with open("../data/all_rebase_targets_250131_6bp_less_deambiguated.txt", "w") as f:
+	for s in unambiguous_sequences:
+		f.write("%s %d\n" % (s, len(s)))
+
+

### 2025-01-31 scripts/Paper-analysis.Rmd
chi-squared test for top depleted motifs

@@
 ```
+
+### All motifs testing for over-representation
+
+Go through values of k for across PTU averages.
+
+```{r}
+# k= 4
+# 6-bp palindromes
+palindromes = read.csv('~/Downloads/trieste/k4.txt', header=F)$V1
+results.df = read.csv('../test_ptus_k4_scores.csv',header=T)
+# Add PTUs as column names
+colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
+# Normalise scores by the number of plasmids within PTU?
+n.plasmids = table(lr.no.duplicates.subset$PTU)
+results.df.norm = results.df %>%
+mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
+# Plot all palindromes
+results.df.long = results.df.norm %>%
+pivot_longer(cols = starts_with("PTU"),
+names_to = "score_type",
+values_to = "score_avg")
+results.df.norm.average = results.df.long %>% group_by(kmer) %>%
+summarise(median=median(score_avg)) %>%
+arrange(median)
+# Take the top 1% depleted
+results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.05))),] # not enough for 0.01
+results.df.norm.average.top.depleted$palindrome = ifelse(results.df.norm.average.top.depleted$kmer %in% palindromes,
+"palindrome",
+"non-palindrome")
+# Read in all RM targets
+all_rm_targets = read.csv('../data/all_rebase_targets_250131_6bp_less_deambiguated.txt', header=F, sep=' ')
+all_rm_targets_4bp =all_rm_targets$V1[all_rm_targets$V2==4]
+results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
+all_rm_targets_4bp,
+"target", "non-target")
+table(results.df.norm.average.top.depleted$palindrome)
+table(results.df.norm.average.top.depleted$target)
+chisq.test(matrix(c(3, 10, 13, 230 ), nrow=2))
+chisq.test(matrix(c(4, 9, 99, 144 ), nrow=2))
+
+### And for k=5
+results.df = read.csv('../test_ptus_k5_scores.csv',header=T)
+# Add PTUs as column names
+colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
+# Normalise scores by the number of plasmids within PTU?
+n.plasmids = table(lr.no.duplicates.subset$PTU)
+results.df.norm = results.df %>%
+mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
+results.df.long = results.df.norm %>%
+pivot_longer(cols = starts_with("PTU"),
+names_to = "score_type",
+values_to = "score_avg")
+results.df.norm.average = results.df.long %>% group_by(kmer) %>%
+summarise(median=median(score_avg)) %>%
+arrange(median)
+# Take the top 1% depleted
+results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),] # Top 5% since so few
+all_rm_targets_5bp =all_rm_targets$V1[all_rm_targets$V2==5]
+results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
+all_rm_targets_5bp,
+"target", "non-target")
+table(results.df.norm.average.top.depleted$target)
+chisq.test(matrix(c(5, 6, 277, 736 ), nrow=2))
+
+# k =6
+# 6-bp palindromes
+palindromes = read.csv('~/Downloads/trieste/k6.txt', header=F)$V1
+results.df = read.csv('../test_ptus_k6_scores.csv',header=T)
+# Add PTUs as column names
+colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
+# Normalise scores by the number of plasmids within PTU?
+n.plasmids = table(lr.no.duplicates.subset$PTU)
+results.df.norm = results.df %>%
+mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
+# Plot all palindromes
+results.df.long = results.df.norm %>%
+pivot_longer(cols = starts_with("PTU"),
+names_to = "score_type",
+values_to = "score_avg")
+results.df.norm.average = results.df.long %>% group_by(kmer) %>%
+summarise(median=median(score_avg)) %>%
+arrange(median)
+# Take the top 1% depleted
+results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),]
+results.df.norm.average.top.depleted$palindrome = ifelse(results.df.norm.average.top.depleted$kmer %in% palindromes,
+"palindrome",
+"non-palindrome")
+# Read in all RM targets
+all_rm_targets = read.csv('../data/all_rebase_targets_250131_6bp_less_deambiguated.txt', header=F, sep=' ')
+all_rm_targets_6bp =all_rm_targets$V1[all_rm_targets$V2==6]
+results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
+all_rm_targets_6bp,
+"target", "non-target")
+table(results.df.norm.average.top.depleted$palindrome)
+table(results.df.norm.average.top.depleted$target)
+chisq.test(matrix(c(7, 34, 865-7, 4096-865-34 ), nrow=2))
+chisq.test(matrix(c(4, 37, 60, 4055 ), nrow=2))
+
+```

### 2025-02-25 scripts/Paper-analysis.Rmd
pangenome statistics plot

@@
 kable(results_df, digits = 3, format = "html")
+```
+
+Make a plot of these results, including information on PTU pangenome size:
+
+```{r}
+ptu_for_plotting = ptu %>% pivot_longer(cols=c("n.core", "n.others"))
+ptu_for_plotting$name = ordered(ptu_for_plotting$name,
+                                levels=c("n.others", "n.core"))
+ggplot(ptu_for_plotting, aes(PTU, value, fill=name))+
+  geom_bar(stat="identity", position="stack")+
+  coord_flip()
+# Want average size of the PTU as well...
+
 ```

### 2025-02-25 scripts/pangenome_statistics.R

@@ -1,0 +1,35 @@
+# Get average size of PTU
+ptu = read.csv('../data/PTU-pangenome-stats.csv')
+for (each_ptu in ptu$PTU){
+  gene_pa_file <- list.files(paste0("/Users/Liam/Downloads/06_plaspan/", each_ptu), 
+                        pattern = "gene_presence_absence.Rtab$", full.names = TRUE, recursive = TRUE)
+  gene_counts = read.csv(gene_pa_file, sep='\t', row.names = 1)
+  ptu$median.genes[which(ptu$PTU==each_ptu)] = median(colSums(gene_counts))
+  ptu$max.genes[which(ptu$PTU==each_ptu)] = max(colSums(gene_counts))
+  ptu$min.genes[which(ptu$PTU==each_ptu)] = min(colSums(gene_counts))
+}
+
+# Plot these
+ptu_for_plotting = ptu %>% pivot_longer(cols=c("n.core", "n.others"))
+
+ptu = ptu[order(ptu$median.genes),]
+ptu$PTU = ordered(ptu$PTU, 
+                  levels=rev(ptu$PTU))
+ptu$leading.region = ifelse(ptu$PTU %in% leading.region.PTUs,
+                            "yes", "no")
+pdf('../results/ptu-summary-pangenome-plot.pdf', width=9, height=5)
+ggplot(ptu[which(ptu$median.genes>20),], aes(PTU, median.genes, ymin=min.genes, ymax=max.genes,
+                                             fill=leading.region))+
+  geom_bar(aes(y=n.core), stat="identity")+
+  geom_errorbar()+
+  geom_point(shape=21)+
+  theme_bw()+
+  xlab("")+
+  ylab("Number of genes")+
+  labs(fill="Leading region analysed")+
+  coord_flip()+
+  scale_color_brewer(palette="Set2")+
+  scale_fill_brewer(palette="Set2")+
+  theme(axis.text=element_text(colour="black"))
+dev.off()
+

### 2025-02-26 scripts/cluster/subset.py
for subsetting fasta files to one rep per family

@@ -1,0 +1,38 @@
+# Subset a pangenome fasta (core/accessory) to one representative per family
+
+import random
+from collections import defaultdict
+import sysi
+
+
+def parse_fasta(filename):
+    """Parse a FASTA file and yield (header, sequence) tuples."""
+    with open(filename, "r") as f:
+        header, seq = None, []
+        for line in f:
+            line = line.strip()
+            if line.startswith(">"):
+                if header:
+                    yield header, "".join(seq)
+                header, seq = line, []
+            else:
+                seq.append(line)
+        if header:
+            yield header, "".join(seq)
+
+def select_representatives(fasta_file, output_file):
+    groups = defaultdict(list)
+
+    # Group sequences by their group ID
+    for header, sequence in parse_fasta(fasta_file):
+        group_id = header.split("|")[-1]  # Extract group ID after '|'
+        groups[group_id].append((header, sequence))
+
+    # Select one random representative per group
+    with open(output_file, "w") as f:
+        for group_id, entries in groups.items():
+            header, sequence = random.choice(entries)
+            f.write(f"{header}\n{sequence}\n")
+
+if __name__=="__main__":
+    select_representatives(sys.argv[1], sys.argv[2])

### 2025-02-26 scripts/run-rmes-score-discrepancy.py
add old script back

@@ -1,0 +1,52 @@
+import subprocess
+import os
+import glob
+
+def run_comparisons(base_script, directory, kmers_file, output_file, k):
+    # Find all core and accessory files in the given directory
+    core_files = sorted(glob.glob(os.path.join(directory, "PTU-*_core_genes_k"+str(k)+".csv")))
+    accessory_files = sorted(glob.glob(os.path.join(directory, "PTU-*_accessory_genes_k"+str(k)+".csv")))
+    if kmers_file=="per-PTU":
+        kmers_files = sorted(glob.glob(os.path.join(directory, "PTU-*_targets.txt")))
+    # Ensure matching pairs
+    if len(core_files) != len(accessory_files):
+        print("Error: Mismatched number of core and accessory files.")
+        return
+
+    with open(output_file, "w") as out:
+        out.write("PTU,n_selected_kmers,discrepancy_selected,discrepancy_others,wilcoxon_stat,p_value\n")
+        for core_file, accessory_file, kmers_file in zip(core_files, accessory_files, kmers_files):
+                # Construct the command
+                command = [
+                    "python", base_script,
+                    "--core", core_file,
+                    "--accessory", accessory_file,
+                    "--kmers", kmers_file
+                ]
+
+                # Execute the command
+                result = subprocess.run(command, capture_output=True, text=True)
+
+                # Output the results
+                if result.returncode == 0:
+                    out.write(result.stdout)
+                else:
+                    print(f"Error for core: {core_file} and accessory: {accessory_file}")
+                    print(result.stderr)
+
+if __name__ == "__main__":
+    import argparse
+
+    # Argument parsing
+    parser = argparse.ArgumentParser(description="Run comparisons for multiple core and accessory files.")
+    parser.add_argument("--script", required=True, help="Path to the comparison script.")
+    parser.add_argument("--directory", required=True, help="Directory containing core and accessory files.")
+    parser.add_argument("--kmers", required=True, help="Path to the kmers text file (make sure it matches value of k). If 'per-PTU' is passed then will use targets by PTU, generated by scripts/create-PTU-target-DB.py")
+    parser.add_argument("--output", required=True, help="Path to the output file.")
+    parser.add_argument("--k", required=True, help="Value of k (4,5,6).")
+
+
+    args = parser.parse_args()
+
+    # Run comparisons
+    run_comparisons(args.script, args.directory, args.kmers, args.output, args.k)

### 2025-02-27 scripts/Paper-analysis.Rmd
one rep per gene family analysis

@@
 ```
+
+### All hard shell vs. accessory, one representative per gene family
+
+We subsampled to one representative per gene family and the same number of hard shell and accessory genes, whichever was smaller (PTU-Bac37 was the only one that had more hard shell than accessory genes).
+```{r}
+core_k6 = read.csv('~/Downloads/trieste/one_per_family/all_core_k6.csv')
+accessory_k6 = read.csv('~/Downloads/trieste/one_per_family/all_accessory_k6.csv')
+merge_k6 = merge(core_k6, accessory_k6, by="word")
+merge_k6$diff = merge_k6$score.y-merge_k6$score.y
+merge_k6$palindrome = ifelse(merge_k6$word %in% palindromes,
+                             "palindrome", "other")
+ggplot(merge_k6, aes(score.x-score.y))+
+  geom_histogram()+
+  facet_wrap(~palindrome)
+```

### 2025-02-27 scripts/run-rmes-score-discrepancy.py

@@ -19,8 +19,8 @@ def run_comparisons(base_script, directory, kmers_file, output_file, k):
                 # Construct the command
                 command = [
                     "python", base_script,
-                    "--core", core_file,
-                    "--accessory", accessory_file,
+                    "--file1", core_file,
+                    "--file2", accessory_file,
                     "--kmers", kmers_file
                 ]
 

### 2025-03-03 scripts/Paper-analysis.Rmd
latest analysis

--- 2025-01-31 scripts/Paper-analysis.Rmd
@@
 
-```
-```{r}
-ggplot(df_4_filter, aes(x="k=4", wilcoxon_stat, colour=p_value<0.05))+
-  geom_point()
 ```
@@
 kable(results_df, digits = 3, format = "html")
+```
+
+Repeated this analysis with one gene representative per family:
+
+```{r}
+# Read in results for RM targets
+df_4 = read.csv('../test_one_per_family_within_range_targets_k4.csv')
+df_5 = read.csv('../test_one_per_family_within_range_targets_k5.csv')
+df_6 = read.csv('../test_one_per_family_within_range_targets_k6.csv')
+
+# Convert discrepancies to numeric (presence of None has messed this up)
+df_4$discrepancy_selected = as.numeric(df_4$discrepancy_selected)
+df_4$discrepancy_selected[is.na(df_4$discrepancy_selected)] = 0
+df_5$discrepancy_selected = as.numeric(df_5$discrepancy_selected)
+df_5$discrepancy_selected[is.na(df_5$discrepancy_selected)] = 0
+df_6$discrepancy_selected = as.numeric(df_6$discrepancy_selected)
+df_6$discrepancy_selected[is.na(df_6$discrepancy_selected)] = 0
+
+# Median of selected k-mers
+df = t(data.frame(c(median(df_4[which(df_4$n_selected_kmers!=0),"discrepancy_selected"]), 
+        wilcox.test(df_4[which(df_4$n_selected_kmers!=0),"discrepancy_selected"])$statistic,
+        wilcox.test(df_4[which(df_4$n_selected_kmers!=0),"discrepancy_selected"])$p.value),
+        c(median(df_5[which(df_5$n_selected_kmers!=0),"discrepancy_selected"]), 
+        wilcox.test(df_5[which(df_5$n_selected_kmers!=0),"discrepancy_selected"])$statistic,
+        wilcox.test(df_5[which(df_5$n_selected_kmers!=0),"discrepancy_selected"])$p.value),
+        c(median(df_6[which(df_6$n_selected_kmers!=0),"discrepancy_selected"]), 
+        wilcox.test(df_6[which(df_6$n_selected_kmers!=0),"discrepancy_selected"])$statistic,
+        wilcox.test(df_6[which(df_6$n_selected_kmers!=0),"discrepancy_selected"])$p.value)))
+colnames(df) = c("median", "statistic", "p.value")
+print(df)
+
+```
+
+### All hard shell vs. accessory, one representative per gene family
+
+We subsampled to one representative per gene family and the same number of hard shell and accessory genes, whichever was smaller (PTU-Bac37 was the only one that had more hard shell than accessory genes).
+```{r}
+palindromes_k4 = tolower(read.csv('~/Downloads/trieste/k4.txt', header=F)$V1)
+palindromes_k6 = tolower(read.csv('~/Downloads/trieste/k6.txt', header=F)$V1)
+
+core_k6 = read.csv('~/Downloads/trieste/one_per_family/all_core_1745k_k6.csv')
+accessory_k6 = read.csv('~/Downloads/trieste/one_per_family/all_accessory_1745k_k6.csv')
+merge_k6 = merge(core_k6, accessory_k6, by="word")
+merge_k6$score.diff = merge_k6$score.x-merge_k6$score.y
+merge_k6$count.diff = merge_k6$count.x-merge_k6$count.y
+
+merge_k6$palindrome = ifelse(merge_k6$word %in% palindromes_k6,
+                             "palindrome", "other")
+median(merge_k6[which(merge_k6$palindrome=="palindrome"),"score.diff"])
+wilcox.test(merge_k6[which(merge_k6$palindrome=="palindrome"),"score.diff"])
+sum(merge_k6$count.x[which(merge_k6$palindrome=="palindrome")])/sum(merge_k6$count.x) * 1000
+sum(merge_k6$count.y[which(merge_k6$palindrome=="palindrome")])/sum(merge_k6$count.y) * 1000
+
+# And for k=4
+
+core_k4 = read.csv('~/Downloads/trieste/one_per_family/all_core_1745k_k4.csv')
+accessory_k4 = read.csv('~/Downloads/trieste/one_per_family/all_accessory_1745k_k4.csv')
+merge_k4 = merge(core_k4, accessory_k4, by="word")
+merge_k4$score.diff = merge_k4$score.x-merge_k4$score.y
+merge_k4$count.diff = merge_k4$count.x-merge_k4$count.y
+
+merge_k4$palindrome = ifelse(merge_k4$word %in% palindromes_k4,
+                             "palindrome", "other")
+median(merge_k4[which(merge_k4$palindrome=="palindrome"),"score.diff"])
+wilcox.test(merge_k4[which(merge_k4$palindrome=="palindrome"),"score.diff"])
+
+# Sum palindromes
+sum(merge_k4$count.x[which(merge_k4$palindrome=="palindrome")])/sum(merge_k4$count.x) * 1000
+sum(merge_k4$count.y[which(merge_k4$palindrome=="palindrome")])/sum(merge_k4$count.y) * 1000
+
 ```
@@
 ggsave(p.lines.expectation+scale_x_continuous(labels = function(x) x / 1000, limits=c(0,50000)), file='../results/6mer-deviation-expectation-0-50-kb.pdf',
+       width=10, height=6)
+
+```
+```{r}
+# Sliding window for GC content
+p.lines.GC = ggplot(d.reduced.median.plot, aes(x, GC/5000 * 100))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("Palindrome depletion per kb (5kb sliding window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+ggsave(p.lines.GC, file='../results/GC-content.pdf',
+       width=10, height=6)
+
+# And for palindrome density (not expected - actual values)
+p.lines.density = ggplot(d.reduced.median.plot, aes(x, median.original*1000))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("Palindrome density per kb (5kb sliding window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+ggsave(p.lines.density, file='../results/6mer-palindrome-density.pdf',
        width=10, height=6)
@@
 ```{r}
+# Normalise
+PTU.counts = table(plasmid.table$PTU)
+n.plasmids = PTU.counts[leading.region.PTUs]
 # k= 4
@@
 # Add PTUs as column names
-colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
-# Normalise scores by the number of plasmids within PTU?
-n.plasmids = table(lr.no.duplicates.subset$PTU)
+colnames(results.df)[2:ncol(results.df)] = leading.region.PTUs
 results.df.norm = results.df %>%
@@
 # Read in all RM targets
-all_rm_targets = read.csv('../data/all_rebase_targets_250131_6bp_less_deambiguated.txt', header=F, sep=' ')
-all_rm_targets_4bp =all_rm_targets$V1[all_rm_targets$V2==4]
+all_rm_targets = read.csv('../data/rebase_all_4_6bp_targets_250131.csv', header=T)
+all_rm_targets_4bp =all_rm_targets$target[all_rm_targets$length==4]
 results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
@@
 table(results.df.norm.average.top.depleted$target)
-chisq.test(matrix(c(3, 10, 13, 230 ), nrow=2))
-chisq.test(matrix(c(4, 9, 99, 144 ), nrow=2))
+# palindromes
+chisq.test(matrix(c(2, 11, 13, 230 ), nrow=2))
+# Targets
+chisq.test(matrix(c(3, 10, 100, 143 ), nrow=2))
 
@@
 # Take the top 1% depleted
-results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),] # Top 5% since so few
-all_rm_targets_5bp =all_rm_targets$V1[all_rm_targets$V2==5]
+results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),] 
+all_rm_targets_5bp =all_rm_targets$target[all_rm_targets$length==5]
 results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
@@
 table(results.df.norm.average.top.depleted$target)
-chisq.test(matrix(c(5, 6, 277, 736 ), nrow=2))
+chisq.test(matrix(c(5, 11, 277, 731 ), nrow=2))
 
@@
 # Add PTUs as column names
-colnames(results.df)[2:ncol(results.df)] = names(table(lr.no.duplicates.subset$PTU))
-# Normalise scores by the number of plasmids within PTU?
-n.plasmids = table(lr.no.duplicates.subset$PTU)
+colnames(results.df)[2:ncol(results.df)] = leading.region.PTUs
 results.df.norm = results.df %>%
 mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
-# Plot all palindromes
+# Switch to long format
 results.df.long = results.df.norm %>%
@@
 values_to = "score_avg")
+
 results.df.norm.average = results.df.long %>% group_by(kmer) %>%
@@
 # Read in all RM targets
-all_rm_targets = read.csv('../data/all_rebase_targets_250131_6bp_less_deambiguated.txt', header=F, sep=' ')
-all_rm_targets_6bp =all_rm_targets$V1[all_rm_targets$V2==6]
+all_rm_targets = read.csv('../data/rebase_all_4_6bp_targets_250131.csv', header=F, sep=' ')
+all_rm_targets_6bp =all_rm_targets$target[all_rm_targets$length==6]
 results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%

### 2025-03-03 scripts/palindrome-densities.R

@@ -1,0 +1,46 @@
+# Analyse palindrome densities in hard shell vs. accessory
+library(ggplot2)
+palindromes_k4 = tolower(read.csv('~/Downloads/trieste/k4.txt', header=F)$V1)
+palindromes_k6 = tolower(read.csv('~/Downloads/trieste/k6.txt', header=F)$V1)
+
+palindromeDensity = function(rmes_table, k=6){
+  rmes_df = read.csv(rmes_table)
+  if (k==4){
+    palindrome_count = sum(rmes_df[which(rmes_df$word %in% palindromes_k4),"count"])
+  }
+  if (k==6){
+    palindrome_count = sum(rmes_df[which(rmes_df$word %in% palindromes_k6),"count"])
+  }
+  return(palindrome_count/sum(rmes_df$count) * 1000)
+}
+
+# Go through PTUs
+PTUs_of_interest = read.csv('../test_PTUs.txt', header=F)$V1
+results_df_k4 = data.frame()
+results_df_k6 = data.frame()
+for (PTU in PTUs_of_interest){
+  results_df_k6 = rbind(results_df_k6, c(PTU, palindromeDensity(paste0('~/Downloads/trieste/', PTU, '_core_genes_k6.csv')),
+    palindromeDensity(paste0('~/Downloads/trieste/', PTU, '_accessory_genes_k6.csv'))))
+  results_df_k4 = rbind(results_df_k4, c(PTU, palindromeDensity(paste0('~/Downloads/trieste/', PTU, '_core_genes_k4.csv'), 4),
+                                      palindromeDensity(paste0('~/Downloads/trieste/', PTU, '_accessory_genes_k4.csv'), 4)))
+  
+}
+colnames(results_df_k6) = c("PTU", "core", "accessory")
+results_df_k6$core = as.numeric(results_df_k6$core)
+results_df_k6$accessory = as.numeric(results_df_k6$accessory)
+results_df_k6$diff = results_df_k6$core-results_df_k6$accessory
+ggplot(results_df_k6, aes(diff))+
+  geom_boxplot()
+
+colnames(results_df_k4) = c("PTU", "core", "accessory")
+results_df_k4$core = as.numeric(results_df_k4$core)
+results_df_k4$accessory = as.numeric(results_df_k4$accessory)
+results_df_k4$diff = results_df_k4$core-results_df_k4$accessory
+
+median(results_df_k4$diff)
+median(results_df_k6$diff)
[…]

### 2025-04-03 scripts/Paper-analysis.Rmd
updating analysis

@@
   facet_wrap(~PTU, scales="free_x", nrow=3)+
-  ylab("Palindrome depletion per kb (5kb sliding window)")+
+  ylab("GC content (5kb sliding window)")+
   labs(alpha="% of plasmids within PTU")+
@@
   mutate(diff.count=palindrome_count[x=="0"]-mean(palindrome_count),
-         diff.density=diff_density[x=="0"]-mean(diff_density)) %>%
+         diff.density=diff_density[x=="0"]-mean(diff_density),
+         diff.GC=(GC_n[x=="0"]-mean(GC_n))/5000 * 100) %>%
   filter(x=="0")
@@
 ```
+
+Now we want to correlate the GC content difference of the first 5kb of the plasmid with the difference in palindrome density
+
+```{r}
+diff.PTU.averages = diff.leading.region.average %>% group_by(PTU) %>%
+  summarise(mean.GC=mean(diff.GC),
+            GC=mean(GC_n[x=="0"]/5000 * 100),
+            mean.palindrome=mean(palindrome_density),
+            mean.palindrome.diff=mean(diff.density) * 1000,
+            n=length(diff.GC))
+ggplot(diff.PTU.averages, aes(mean.GC, mean.palindrome.diff))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  ggrepel::geom_text_repel(aes(label=PTU))+
+    geom_point(aes(size=n), shape=21, colour="black", fill="grey")+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  xlab("GC difference between first 5kb\nand plasmid average (%)")+
+  ylab("Additional palindrome depletion in leading region (per kb)")
+
+# Calculate actual GC and actual palindrome density
+# for a given x-value. I think 5kb is appropriate for leading region to 
+# get rid of start effects. 
+gc_data_frame = function(position_on_plasmid){
+  d.reduced %>% 
+  group_by(plasmid, PTU) %>%
+  mutate(GC=GC_n[x==position_on_plasmid]/5000 * 100,
+         diff.GC=(GC_n[x==position_on_plasmid]-mean(GC_n))/5000 * 100,
+         density=palindrome_density[x==position_on_plasmid]*1000,
+         density.depletion=diff_density[x==position_on_plasmid]*1000) %>%
+  filter(x==position_on_plasmid)
+}
+d.GC.palindrome = gc_data_frame(5000)
+ggplot(d.GC.palindrome, aes(diff.GC, density.depletion))+
+  geom_point()
+
+d.GC.palindrome.PTU = d.GC.palindrome %>% group_by(PTU) %>%
+  summarise(GC=mean(GC),
+            diff.GC=mean(diff.GC),
+            density=mean(density),
+            density.depletion=mean(density.depletion),
+            n=length(window_size))
+ggplot(d.GC.palindrome, aes(diff.GC, density.depletion, fill=PTU))+
+  geom_point(shape=21, colour="black")+
+  facet_wrap(~PTU)
+p.gc.correlation = ggplot(d.GC.palindrome, aes(diff.GC, density.depletion))+
+  geom_point()+
+  theme_bw()+
+  xlab("Relative GC content at 5-10kb (%)")+
+  ylab("Relative palindrome depletion at 5-10kb (per kb)")+
+    xlim(c(-15, 10))+
+  ylim(-13, 14)
+p.gc.correlation.ptu =  ggplot(d.GC.palindrome.PTU, aes(diff.GC, density.depletion))+
+  geom_point(aes(size=n, fill=PTU), shape=21, colour="black")+
+  theme_bw()+
+  ggrepel::geom_text_repel(aes(label=PTU), size=2, min.segment.length = 0)+
+  guides(fill=FALSE)+
+    xlab("Relative GC content (%)")+
+  ylab("Relative palindrome depletion (per kb)")
+pdf('../results/GC-correlation-depletion.pdf', width=8, height=4)
+cowplot::plot_grid(p.gc.correlation+ggtitle("(a)"), p.gc.correlation.ptu+ggtitle("(b)"))
+dev.off()
+cor.test(d.GC.palindrome$diff.GC, d.GC.palindrome$density.depletion, method="spearman")
+cor.test(d.GC.palindrome.PTU$diff.GC, d.GC.palindrome.PTU$density.depletion)
+
+summary(lm(density.depletion ~ diff.GC*PTU, data=d.GC.palindrome))
+
+```
+
 ### Averaging over all plasmids
@@
        width=3, height=3)
+```
+Also do this for GC content.
+
+```{r}
+overall.median.GC = d.reduced %>% filter(x<50000) %>%
+  group_by(x) %>%
+  summarise(y=mean(GC_n),
+            n=length(GC_n))
+p.overall.average.GC = ggplot(overall.median.GC, aes(x, y/5000 * 100))+
+  geom_line()+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  geom_hline(yintercept = 50, linetype='dashed')+
+  ylab("GC content (per kb)")+
+    scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(axis.text=element_text(colour="black"))
+p.overall.average.GC
+# Save
+ggsave(p.overall.average.GC, file='../results/GC-mean-overall.pdf', 
+       width=5, height=2)
 ```

### 2025-04-03 scripts/Paper-analysis.Rmd
link figure titles to code

@@
        width=10, height=6)
-# Plot only first 25kb
+# Plot only first 50kb
 ggsave(p.lines.expectation+scale_x_continuous(labels = function(x) x / 1000, limits=c(0,50000)), file='../results/6mer-deviation-expectation-0-50-kb.pdf',
@@
 
-### Compare leading region to rest of plasmid
+### Figure 2
+
+Figure 2. RM target depletion in the leading region of conjugative plasmids.
+
+```{r}
+overall.median = d.reduced %>% filter(x<50000) %>%
+  group_by(x) %>%
+  summarise(y=median(diff_density),
+            n=length(diff_density))
+p.overall.average = ggplot(overall.median, aes(x, y*1000))+
+  geom_line()+
+  scale_y_continuous(breaks=seq(-10, 0, 2), limits = c(-9, 0))+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  ylab("Palindrome depletion (per kb)")+
+    scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(axis.text=element_text(colour="black"))
+p.overall.average
+# Save
+ggsave(p.overall.average, file='../results/Figure-2-6mer-deviation-overall.pdf', 
+       width=3, height=3)
+```
 
@@
   scale_y_continuous(breaks=seq(-8, 4, 2), limits=c(-8,5))
-ggsave(p.leading.region.diff.density, file='../results/6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
+ggsave(p.leading.region.diff.density, file='../results/Figure-2-6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
        width=5.5, height=6)
@@
 
-Now we want to correlate the GC content difference of the first 5kb of the plasmid with the difference in palindrome density
+### Figure 3
+
+Figure 3.
+GC-rich leading regions are more depleted in 6-bp palindromes.
+Values are calculated for the 5-10kb section of the leading region relative to the mean average value over the whole length of the plasmid.
+Spearman’s correlations are given as inset values.
+The relationship between relative GC content and palindrome depletion was still significant in a linear model controlling for PTU (depletion ~ GC*PTU).
 
@@
   ylab("Relative palindrome depletion (per kb)")
-pdf('../results/GC-correlation-depletion.pdf', width=8, height=4)
+pdf('../results/Figure-3-GC-correlation-depletion.pdf', width=8, height=4)
 cowplot::plot_grid(p.gc.correlation+ggtitle("(a)"), p.gc.correlation.ptu+ggtitle("(b)"))
@@
 
-Similar to Samuel et al.'s display of anti-defence gene density, average over all conjugative plasmids.
-
-```{r}
-overall.median = d.reduced %>% filter(x<50000) %>%
-  group_by(x) %>%
-  summarise(y=median(diff_density),
-            n=length(diff_density))
-p.overall.average = ggplot(overall.median, aes(x, y*1000))+
-  geom_line()+
-  scale_y_continuous(breaks=seq(-10, 0, 2), limits = c(-9, 0))+
-  theme_bw()+
-  theme(panel.grid = element_blank())+
-  geom_hline(yintercept = 0, linetype='dashed')+
-  ylab("Palindrome depletion (per kb)")+
-    scale_x_continuous(labels = function(x) x / 1000) +
-  xlab("Position on plasmid (kb)")+
-  theme(axis.text=element_text(colour="black"))
-p.overall.average
-# Save
-ggsave(p.overall.average, file='../results/6mer-deviation-overall.pdf', 
-       width=3, height=3)
-```
 Also do this for GC content.
@@
 
-Test for a particular PTU.
+### Figure 4
 
+Figure 4: Average depletion of within-range RM targets in the leading vs. lagging regions for conjugative PTUs. 
 ```{r}
@@
   theme(axis.text.y=element_text(hjust=0))
-ggsave(p.depletion.targets, file='../results/figure-depletion-within-range-targets.pdf',
+ggsave(p.depletion.targets, file='../results/Figure-4-figure-depletion-within-range-targets.pdf',
        width=10, height=6)

### 2025-04-03 scripts/Paper-analysis.Rmd
add Figure 1 into main script

@@
 library(tidyr)
+```
+### Figure 1
+
+Figure 1. Summary of the 36 plasmid taxonomic units (PTUs) analysed in this paper.
+
+```{r}
+# Get average size of PTU
+ptu = read.csv('../data/PTU-pangenome-stats.csv')
+for (each_ptu in ptu$PTU){
+  gene_pa_file <- list.files(paste0("/Users/Liam/Downloads/06_plaspan/", each_ptu), 
+                        pattern = "gene_presence_absence.Rtab$", full.names = TRUE, recursive = TRUE)
+  gene_counts = read.csv(gene_pa_file, sep='\t', row.names = 1)
+  ptu$median.genes[which(ptu$PTU==each_ptu)] = median(colSums(gene_counts))
+  ptu$max.genes[which(ptu$PTU==each_ptu)] = max(colSums(gene_counts))
+  ptu$min.genes[which(ptu$PTU==each_ptu)] = min(colSums(gene_counts))
+}
+
+# Plot these
+ptu_for_plotting = ptu %>% pivot_longer(cols=c("n.core", "n.others"))
+
+ptu = ptu[order(ptu$median.genes),]
+ptu$PTU = ordered(ptu$PTU, 
+                  levels=rev(ptu$PTU))
+ptu$leading.region = ifelse(ptu$PTU %in% leading.region.PTUs,
+                            "yes", "no")
+pdf('../results/Figure-1-ptu-summary-pangenome-plot.pdf', width=9, height=5)
+ggplot(ptu[which(ptu$median.genes>20),], aes(PTU, median.genes, ymin=min.genes, ymax=max.genes,
+                                             fill=leading.region))+
+  geom_bar(aes(y=n.core), stat="identity")+
+  geom_errorbar()+
+  geom_point(shape=21)+
+  theme_bw()+
+  xlab("")+
+  ylab("Number of genes")+
+  labs(fill="Leading region analysed")+
+  coord_flip()+
+  scale_color_brewer(palette="Set2")+
+  scale_fill_brewer(palette="Set2")+
+  theme(axis.text=element_text(colour="black"))
+dev.off()
+
 ```
@@
 # Pangenome statistics for PTUs
-ptu = read.csv('../data/PTU-pangenome-stats.csv')
+#ptu = read.csv('../data/PTU-pangenome-stats.csv')
 
@@
   theme(legend.position = c(0.8, 0.2))
-ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf',
+ggsave(p.lines.expectation, file='../results/Figure-S2-6mer-deviation-expectation.pdf',
        width=10, height=6)
@@
   theme(legend.position = c(0.8, 0.2))
-ggsave(p.lines.density, file='../results/6mer-palindrome-density.pdf',
+ggsave(p.lines.density, file='../results/Figure-S3-6mer-palindrome-density.pdf',
        width=10, height=6)
@@
 wilcox.test(merge_scores_counts$score_avg)
-ggsave(scores.plot, file='../results/6mer-palindromes-scores.pdf',
+ggsave(scores.plot, file='../results/Figure-S4-6mer-palindromes-scores.pdf',
        height=10, width=10)

### 2025-04-03 scripts/Paper-analysis.Rmd
figure names

@@
 library(tidyr)
+```
+
+Some preliminary statistics about discarded PTUs (for main text).
+
+```{r}
+lr = read.csv('../data/leading_region_081123.tsv', sep='\t')
+plasmid.table = read.csv('../data/plasmid_table.tsv', sep='\t')
+retained.PTUs.lr = read.csv('../data/leading-region-PTUs.txt', header=F)$V1
+lr.identified.PTU.names = sort(unique(lr$PTU))
+# Minimum sizes
+PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
+  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
+```
+
+We discard those where the minimum size is <20kb: E1, E3, E9, E10, E14, E76. We check the maximum size per PTU:
+
+```{r}
+sort(sapply(paste0("PTU-E", c("1", "3", "9", "10", "14", "76")), function(x) 
+  max(plasmid.table[plasmid.table$PTU==x, "Size"])))
+```
+
+These PTUs are retained for analysing the leading region:
+
+```{r}
+leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
+cat(leading.region.PTUs, sep="\n")
+```
+
+For convenience for other scripts, I save some files with names of plasmids within each PTU.
+
+```{r}
+for (PTU in unique(plasmid.table$PTU)){
+  cat(plasmid.table$ID[plasmid.table$PTU==PTU], sep="\n",
+      file=paste0('~/Downloads/trieste/', PTU, '_plasmids.txt'))
+}
+
 ```
@@
 
-Some preliminary statistics about discarded PTUs (for main text).
-
-```{r}
-lr = read.csv('../data/leading_region_081123.tsv', sep='\t')
-plasmid.table = read.csv('../data/plasmid_table.tsv', sep='\t')
-retained.PTUs.lr = read.csv('../data/leading-region-PTUs.txt', header=F)$V1
-lr.identified.PTU.names = sort(unique(lr$PTU))
-# Minimum sizes
-PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
-  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
-```
-
-We discard those where the minimum size is <20kb: E1, E3, E9, E10, E14, E76. We check the maximum size per PTU:
-
-```{r}
-sort(sapply(paste0("PTU-E", c("1", "3", "9", "10", "14", "76")), function(x) 
-  max(plasmid.table[plasmid.table$PTU==x, "Size"])))
-```
-
-These PTUs are retained for analysing the leading region:
-
-```{r}
-leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
-cat(leading.region.PTUs, sep="\n")
-```
-
-For convenience for other scripts, I save some files with names of plasmids within each PTU.
-
-```{r}
-for (PTU in unique(plasmid.table$PTU)){
-  cat(plasmid.table$ID[plasmid.table$PTU==PTU], sep="\n",
-      file=paste0('~/Downloads/trieste/', PTU, '_plasmids.txt'))
-}
-
-```
 ### Sliding window analysis
@@
   theme(legend.position = c(0.8, 0.2))
-ggsave(p.lines.GC, file='../results/GC-content.pdf',
+ggsave(p.lines.GC, file='../results/Figure-S1-GC-content.pdf',
        width=10, height=6)
@@
 # Save
-ggsave(p.overall.average.GC, file='../results/GC-mean-overall.pdf', 
+ggsave(p.overall.average.GC, file='../results/Figure-S1-GC-mean-overall.pdf', 
        width=5, height=2)

### 2025-04-03 scripts/pangenome_statistics.R

@@ -1,35 +1,0 @@
-# Get average size of PTU
-ptu = read.csv('../data/PTU-pangenome-stats.csv')
-for (each_ptu in ptu$PTU){
-  gene_pa_file <- list.files(paste0("/Users/Liam/Downloads/06_plaspan/", each_ptu), 
-                        pattern = "gene_presence_absence.Rtab$", full.names = TRUE, recursive = TRUE)
-  gene_counts = read.csv(gene_pa_file, sep='\t', row.names = 1)
-  ptu$median.genes[which(ptu$PTU==each_ptu)] = median(colSums(gene_counts))
-  ptu$max.genes[which(ptu$PTU==each_ptu)] = max(colSums(gene_counts))
-  ptu$min.genes[which(ptu$PTU==each_ptu)] = min(colSums(gene_counts))
-}
-
-# Plot these
-ptu_for_plotting = ptu %>% pivot_longer(cols=c("n.core", "n.others"))
-
-ptu = ptu[order(ptu$median.genes),]
-ptu$PTU = ordered(ptu$PTU, 
-                  levels=rev(ptu$PTU))
-ptu$leading.region = ifelse(ptu$PTU %in% leading.region.PTUs,
-                            "yes", "no")
-pdf('../results/ptu-summary-pangenome-plot.pdf', width=9, height=5)
-ggplot(ptu[which(ptu$median.genes>20),], aes(PTU, median.genes, ymin=min.genes, ymax=max.genes,
-                                             fill=leading.region))+
-  geom_bar(aes(y=n.core), stat="identity")+
-  geom_errorbar()+
-  geom_point(shape=21)+
-  theme_bw()+
-  xlab("")+
-  ylab("Number of genes")+
-  labs(fill="Leading region analysed")+
-  coord_flip()+
-  scale_color_brewer(palette="Set2")+
-  scale_fill_brewer(palette="Set2")+
-  theme(axis.text=element_text(colour="black"))
-dev.off()
-

### 2025-04-17 scripts/Paper-analysis.Rmd
cleaning up Paper-analysis.Rmd and outputs

--- 2025-04-03 scripts/Paper-analysis.Rmd
@@
-```{r}
+---
+title: "Shaw et al. (2025) Supplementary analysis"
+author: "Liam Shaw"
+output:
+  html_document:
+    toc: true
+---
+
+### Preliminaries
+
+This script contains analysis and figure generation for Shaw et al.
+(2025), 'The leading region of many conjugative plasmids is depleted in restriction-modification targets'.
+For more details, see the associated paper.
+The code matches the order of appearance of the figures/analysis in the paper and is separated into sections based on the figures.
+Figure S5 and S6 are not included in this code.
+
+It is intended to be run from inside the `scripts` directory of this repository. Results are saved to to `results` within the repository.
+
+**Note** This relies on large downloaded files and is not intended to work out-of-the-box.
+For example, Roary gene presence/absence tables on my computer are stored in `/Users/Liam/Downloads/06_plaspan/`.
+
+## Dataset
+
+### Statistics
+
+First, we load libraries.
+
+```{r, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(warning = FALSE)
 library(ggplot2)
@@
 
-## Compare hard shell genes vs. accessory
+We calculate some preliminary statistics about the PTUs from our dataset of 50 PTUs that we analyse (see main text).
+
+```{r}
+lr = read.csv('../data/leading_region_081123.tsv', sep='\t')
+plasmid.table = read.csv('../data/plasmid_table.tsv', sep='\t')
+retained.PTUs.lr = read.csv('../data/leading-region-PTUs.txt', header=F)$V1
+lr.identified.PTU.names = sort(unique(lr$PTU))
+# Minimum sizes
+PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
+  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
+```
+
+We discard those where the minimum size of plasmids in the PTU is <20kb: E1, E3, E9, E10, E14, E76. We check the maximum size per PTU:
+
+```{r}
+sort(sapply(paste0("PTU-E", c("1", "3", "9", "10", "14", "76")), function(x) 
+  max(plasmid.table[plasmid.table$PTU==x, "Size"])))
+```
+
+These PTUs are retained for analysing the leading region:
+
+```{r}
+leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
+cat(leading.region.PTUs, sep="\n")
+```
+
+For convenience for other scripts, we save some files with names of plasmids within each PTU.
+
+```{r}
+for (PTU in unique(plasmid.table$PTU)){
+  cat(plasmid.table$ID[plasmid.table$PTU==PTU], sep="\n",
+      file=paste0('~/Downloads/trieste/', PTU, '_plasmids.txt'))
+}
+
+```
+
+### Figure 1
+
+Figure 1. Summary of the 36 plasmid taxonomic units (PTUs) analysed in this paper.
+
+```{r, message=FALSE}
+ptu = read.csv('../data/PTU-pangenome-stats.csv')
+for (each_ptu in ptu$PTU){
+  gene_pa_file <- list.files(paste0("/Users/Liam/Downloads/06_plaspan/", each_ptu), 
+                        pattern = "gene_presence_absence.Rtab$", full.names = TRUE, recursive = TRUE)
+  gene_counts = read.csv(gene_pa_file, sep='\t', row.names = 1)
+  ptu$median.genes[which(ptu$PTU==each_ptu)] = median(colSums(gene_counts))
+  ptu$max.genes[which(ptu$PTU==each_ptu)] = max(colSums(gene_counts))
+  ptu$min.genes[which(ptu$PTU==each_ptu)] = min(colSums(gene_counts))
+}
+
+# Set up for plotting
+ptu_for_plotting = ptu %>% pivot_longer(cols=c("n.core", "n.others"))
+# Order by median size (no. of genes)
+ptu = ptu[order(ptu$median.genes),]
+ptu$PTU = ordered(ptu$PTU, 
+                  levels=rev(ptu$PTU))
+ptu$leading.region = ifelse(ptu$PTU %in% leading.region.PTUs,
+                            "yes", "no")
+# Make the figure
+pdf('../results/Figure-1-ptu-summary-pangenome-plot.pdf', width=9, height=5)
+p.figure1 = ggplot(ptu[which(ptu$median.genes>20),], aes(PTU, median.genes, ymin=min.genes, ymax=max.genes,
+                                             fill=leading.region))+
+  geom_bar(aes(y=n.core), stat="identity")+
+  geom_errorbar()+
+  geom_point(shape=21)+
+  theme_bw()+
+  xlab("")+
+  ylab("Number of genes")+
+  labs(fill="Leading region analysed")+
+  coord_flip()+
+  scale_color_brewer(palette="Set2")+
+  scale_fill_brewer(palette="Set2")+
+  theme(axis.text=element_text(colour="black"))
+p.figure1
+dev.off()
+p.figure1
+```
+
+## Sliding window analysis
+
+### Figures S1-S3
+
+We plot results of sliding window analysis over the conjugative plasmids for k=6. This produces Figure
+
+```{r, message=FALSE}
+# Theoretical expectation for 6-bp palindrome density
+theoreticalPalindromeExpectation = function(p){# p is GC content
+  return ((p**2 - p + 1/2)**3)
+}
+d = read.csv('../test_output_window5000_step500.csv', header=F, stringsAsFactors = F)
+colnames(d) = c("PTU", "plasmid", "x", "window_size", "palindrome_count", "palindrome_density", "GC_n")
+
+d.reduced = d[which(d$PTU %in% leading.region.PTUs), ]
+
+d.reduced$expected_density = theoreticalPalindromeExpectation(d.reduced$GC_n/d.reduced$window_size)
+
+d.reduced$diff_density = d.reduced$palindrome_density-d.reduced$expected_density
+
+# One line per plasmid
+d.reduced.median = d.reduced %>% group_by(PTU, x) %>% 
+  summarise(median=median(diff_density),
+            n=length(diff_density), 
+            median.original=median(palindrome_density),
+            GC=median(GC_n))
+# Maximum counts
+d.reduced.median.plot = d.reduced.median %>% 
+  group_by(PTU) %>%
+  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
+         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
+  ungroup() %>%  # Remove the grouping
+  select(-max_n)  # Optionally remove the max_n column if not needed
+
+# Sliding window for GC content, S1
+p.figureS1 = ggplot(d.reduced.median.plot, aes(x, GC/5000 * 100))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("GC content (5kb sliding window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+ggsave(p.figureS1, file='../results/Figure-S1-GC-content.pdf',
+       width=10, height=6)
+p.figureS1
+
+overall.median.GC = d.reduced %>% filter(x<50000) %>%
+  group_by(x) %>%
+  summarise(y=mean(GC_n),
+            n=length(GC_n))
+p.figureS1.overall = ggplot(overall.median.GC, aes(x, y/5000 * 100))+
+  geom_line()+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  geom_hline(yintercept = 50, linetype='dashed')+
+  ylab("GC content (per kb)")+
+    scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(axis.text=element_text(colour="black"))
+p.figureS1.overall
+# Save
+ggsave(p.figureS1.overall, file='../results/Figure-S1-GC-mean-overall.pdf', 
+       width=5, height=2)
+
+# Plot the distribution, S2
+p.figureS2 = ggplot(d.reduced.median.plot, aes(x, median*1000))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("Palindrome depletion per kb (5kb sliding window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+ggsave(p.figureS2, file='../results/Figure-S2-6mer-deviation-expectation.pdf',
+       width=10, height=6)
+p.figureS2
+```
+
+```{r}
+# And for palindrome density (not expected - actual values)
+p.figureS3 = ggplot(d.reduced.median.plot, aes(x, median.original*1000))+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  theme_bw()+
+  geom_path(aes(group=PTU, alpha=normalized_n))+
+  facet_wrap(~PTU, scales="free_x", nrow=3)+
+  ylab("Palindrome density per kb (5kb sliding window)")+
+  labs(alpha="% of plasmids within PTU")+
+  theme(panel.grid=element_blank())+  
+  scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(legend.position = c(0.8, 0.2))
+ggsave(p.figureS3, file='../results/Figure-S3-6mer-palindrome-density.pdf',
+       width=10, height=6)
+p.figureS3
+```
+
+### Figure 2
+
+Figure 2. RM target depletion in the leading region of conjugative plasmids.
+
+```{r}
+overall.median = d.reduced %>% filter(x<50000) %>%
+  group_by(x) %>%
+  summarise(y=median(diff_density),
+            n=length(diff_density))
+p.overall.average = ggplot(overall.median, aes(x, y*1000))+
+  geom_line()+
+  scale_y_continuous(breaks=seq(-10, 0, 2), limits = c(-9, 0))+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  geom_hline(yintercept = 0, linetype='dashed')+
+  ylab("Palindrome depletion (per kb)")+
+    scale_x_continuous(labels = function(x) x / 1000) +
+  xlab("Position on plasmid (kb)")+
+  theme(axis.text=element_text(colour="black"))
+p.overall.average
+# Save
+ggsave(p.overall.average, file='../results/Figure-2-6mer-deviation-overall.pdf', 
+       width=3, height=3)
+```
+
+This is the main panel of the figure.
+
+```{r}
+# Start from d.reduced, which contains per-plasmid results with 5000bp sliding window and step size of 500bp
+# Compare depletion in first 5kb with average for the rest of the plasmid
+diff.leading.region.average = d.reduced %>% 
+  group_by(plasmid, PTU) %>%
+  mutate(diff.count=palindrome_count[x=="0"]-mean(palindrome_count),
+         diff.density=diff_density[x=="0"]-mean(diff_density),
+         diff.GC=(GC_n[x=="0"]-mean(GC_n))/5000 * 100) %>%
+  filter(x=="0")
+
+# Order the PTUs by mean values of depletion for better visual display
+mean.values = diff.leading.region.average %>% group_by(PTU) %>%
+  summarise(mean=mean(diff.density)*1000)
+diff.leading.region.average$PTU = ordered(gsub("PTU-", "", diff.leading.region.average$PTU),
+                                          levels=gsub("PTU-", "", mean.values$PTU[order(mean.values$mean)]))
+p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
+  ggbeeswarm::geom_quasirandom(alpha=0.8)+
+  theme_bw()+
+  theme(panel.grid = element_blank())+
+  geom_hline(yintercept = 0)+
+  xlab("")+
+  theme(axis.text=element_text(colour="black", size=18))+
+  ylab("Palindrome depletion in leading region (per kb)")+
+  theme(axis.text.x=element_text(angle=45, hjust=1))+
+  ylab("")
+p.leading.region.diff.density
+# Based on this, remove two outliers
+p.leading.region.diff.density = p.leading.region.diff.density+
+  scale_y_continuous(breaks=seq(-8, 4, 2), limits=c(-8,5))
+ggsave(p.leading.region.diff.density, file='../results/Figure-2-6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
+       width=5.5, height=6)
+```
+
+This is the inset panel which is added into the panel above to create the figure in the paper.
+
+### Figure 3
+
+Figure 3.
+GC-rich leading regions are more depleted in 6-bp palindromes.
+Values are calculated for the 5-10kb section of the leading region relative to the mean average value over the whole length of the plasmid.
+Spearman’s correlations are given as inset values.
+The relationship between relative GC content and palindrome depletion was still significant in a linear model controlling for PTU (depletion ~ GC*PTU).
+
+```{r}
+diff.PTU.averages = diff.leading.region.average %>% group_by(PTU) %>%
+  summarise(mean.GC=mean(diff.GC),
+            GC=mean(GC_n[x=="0"]/5000 * 100),
+            mean.palindrome=mean(palindrome_density),
+            mean.palindrome.diff=mean(diff.density) * 1000,
+            n=length(diff.GC))
+
+# Calculate actual GC and actual palindrome density
+# for a given x-value. 5kb is appropriate for leading region to 
+# get the strongest signal (see Figure 2). 
+gc_data_frame = function(position_on_plasmid){
+  d.reduced %>% 
+  group_by(plasmid, PTU) %>%
+  mutate(GC=GC_n[x==position_on_plasmid]/5000 * 100,
+         diff.GC=(GC_n[x==position_on_plasmid]-mean(GC_n))/5000 * 100,
+         density=palindrome_density[x==position_on_plasmid]*1000,
+         density.depletion=diff_density[x==position_on_plasmid]*1000) %>%
+  filter(x==position_on_plasmid)
+}
+d.GC.palindrome = gc_data_frame(5000)
+
+d.GC.palindrome.PTU = d.GC.palindrome %>% group_by(PTU) %>%
+  summarise(GC=mean(GC),
+            diff.GC=mean(diff.GC),
+            density=mean(density),
+            density.depletion=mean(density.depletion),
+            n=length(window_size))
+
+p.figure3a = ggplot(d.GC.palindrome, aes(diff.GC, density.depletion))+
+  geom_point()+
+  theme_bw()+
+  xlab("Relative GC content at 5-10kb (%)")+
+  ylab("Relative palindrome depletion at 5-10kb (per kb)")+
+    xlim(c(-15, 10))+
+  ylim(-13, 14)
+p.figure3b =  ggplot(d.GC.palindrome.PTU, aes(diff.GC, density.depletion))+
+  geom_point(aes(size=n, fill=PTU), shape=21, colour="black")+
+  theme_bw()+
+  ggrepel::geom_text_repel(aes(label=PTU), size=2, min.segment.length = 0)+
+  guides(fill=FALSE)+
+    xlab("Relative GC content (%)")+
+  ylab("Relative palindrome depletion (per kb)")
+pdf('../results/Figure-3-GC-correlation-depletion.pdf', width=8, height=4)
+p.figure3 = cowplot::plot_grid(p.figure3a+ggtitle("(a)"), p.figure3b+ggtitle("(b)"))
+p.figure3
+dev.off()
+p.figure3
+# Calculate correlations
+cor.test(d.GC.palindrome$diff.GC, d.GC.palindrome$density.depletion, method="spearman")
+cor.test(d.GC.palindrome.PTU$diff.GC, d.GC.palindrome.PTU$density.depletion)
+```
+
+## Depletion of specific targets
+
+### Figure 4
+
+Figure 4: Average depletion of within-range RM targets in the leading vs. lagging regions for conjugative PTUs.
+
+```{r}
+
+results = NULL
+for (PTU in leading.region.PTUs){
+  for (k in c(5,6)){
+    scores_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_scores.csv'), header=T)
+    n_targets = nrow(scores_df)
+    scores_average <- scores_df %>%
+  rowwise() %>%
+  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
+  ungroup() %>%
+  select(kmer, score_avg)
+    counts_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_counts.csv'), header=T)
+    counts_average <- counts_df %>%
+  rowwise() %>%
+  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
+  ungroup() %>%
+  select(kmer, count_avg)
+    results = rbind(results, c(PTU, k, n_targets, median(scores_average$score_avg), median(counts_average$count_avg)))
+  }
+}
+
+results = data.frame(results)
+colnames(results) = c("PTU", "k", "n.targets", "median.score", "median.count")
+results$median.score= as.numeric(results$median.score)
+results$median.count = as.numeric(results$median.count)
+
+# Order by median score for k=6
+order.k.6.scores = results[which(results$k==6), "PTU"][order(results[which(results$k==6), "median.score"])]
+results$PTU = ordered(results$PTU,
+                      levels=rev(order.k.6.scores))
+
+p.figure4 = ggplot(results, aes(median.score, as.factor(PTU)))+
+  geom_vline(xintercept = 0, size=1)+
+  geom_bar(stat="identity", position="dodge")+
+  facet_wrap(~k, scales="free_x")+
+  ylab("")+
+  xlab("Median depletion score of within-range RM targets\n(leading vs. lagging, 10kb)")+
+   theme_bw()+
+  theme(panel.grid = element_blank())+
+  theme(axis.text=element_text(colour="black", size=12),
+        axis.title.x=element_text(size=24),
+        axis.text.x=element_text(size=18))+
+  theme(axis.text.y=element_text(hjust=0))
+ggsave(p.figure4, file='../results/Figure-4-figure-depletion-within-range-targets.pdf',
+       width=10, height=6)
+```
+
+### Figure S4
+
+```{r}
+scores_df = read.csv('../test_6mer_leading_lagging_scores.csv', header=T)
+scores_average <- scores_df %>%
+  rowwise() %>%
+  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
+  ungroup() %>%
+  select(kmer, score_avg)  
+# Order by average score
+scores_average$kmer = ordered(scores_average$kmer,
+                              levels=scores_average$kmer[order(scores_average$score_avg)])
+scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
+  coord_flip()
+
+# Same for counts
+counts_df = read.csv('../test_6mer_leading_lagging_counts.csv', header=T)
+counts_average <- counts_df %>%
+  rowwise() %>%
+  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
+  ungroup() %>%
+  select(kmer, count_avg)
+
+# Merge together
+merge_scores_counts = merge(counts_average, scores_average, by="kmer")
+# average score and average count are well-correlated
+ggplot(merge_scores_counts, aes(count_avg, score_avg))+geom_point()
+
+# Order by average score
+merge_scores_counts$kmer = ordered(merge_scores_counts$kmer,
+                              levels=scores_average$kmer[order(scores_average$score_avg)])
+p.figureS4 = ggplot(merge_scores_counts, aes(score_avg, kmer))+
+    geom_vline(xintercept = 0)+
+  geom_bar(stat="identity")+
+  geom_vline(xintercept = median(merge_scores_counts$score_avg), linetype='dashed', colour='red')+
+  theme_bw()+
+  theme(axis.text=element_text(colour="black", size=12),
+        axis.title.x=element_text(size=24),
+        axis.text.x=element_text(size=18))+
+  theme(axis.text.y=element_text(hjust=0))+
+  ylab("")+
+  xlab("Mean depletion score\n(leading vs. lagging, 10kb)")
+wilcox.test(merge_scores_counts$score_avg)
+ggsave(p.figureS4, file='../results/Figure-S4-6mer-palindromes-scores.pdf',
+       height=10, width=10)
+```
+
+## Other analysis
+
+For more details on the analysis below, see Supplementary Text 1.
 
@@
 
-Repeated this analysis with one gene representative per family:
+We repeat this analysis with one gene representative per family:
 
@@
 print(df)
-
 ```
 
-### All hard shell vs. accessory, one representative per gene family
+We then subsample to one representative per gene family and the same number of hard shell and accessory genes, whichever is smaller (PTU-Bac37 is the only PTU that had more hard shell than accessory genes).
 
-We subsampled to one representative per gene family and the same number of hard shell and accessory genes, whichever was smaller (PTU-Bac37 was the only one that had more hard shell than accessory genes).
 ```{r}
@@
 
-```
-
-## Analysis of leading region
-
-Some preliminary statistics about discarded PTUs (for main text).
-
-```{r}
-lr = read.csv('../data/leading_region_081123.tsv', sep='\t')
-plasmid.table = read.csv('../data/plasmid_table.tsv', sep='\t')
-retained.PTUs.lr = read.csv('../data/leading-region-PTUs.txt', header=F)$V1
-lr.identified.PTU.names = sort(unique(lr$PTU))
-# Minimum sizes
-PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
-  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
-```
-
-We discard those where the minimum size is <20kb: E1, E3, E9, E10, E14, E76. We check the maximum size per PTU:
-
-```{r}
-sort(sapply(paste0("PTU-E", c("1", "3", "9", "10", "14", "76")), function(x) 
-  max(plasmid.table[plasmid.table$PTU==x, "Size"])))
-```
-
-These PTUs are retained for analysing the leading region:
-
-```{r}
-leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
-cat(leading.region.PTUs, sep="\n")
-```
-
-For convenience for other scripts, I save some files with names of plasmids within each PTU.
-
-```{r}
-for (PTU in unique(plasmid.table$PTU)){
-  cat(plasmid.table$ID[plasmid.table$PTU==PTU], sep="\n",
-      file=paste0('~/Downloads/trieste/', PTU, '_plasmids.txt'))
-}
-
-```
-### Sliding window analysis
-
-k=6 
-```{r}
-# Theoretical expectation for 6-bp palindrome density
-theoreticalPalindromeExpectation = function(p){# p is GC content
-  return ((p**2 - p + 1/2)**3)
-}
-d = read.csv('../test_output_window5000_step500.csv', header=F, stringsAsFactors = F)
-colnames(d) = c("PTU", "plasmid", "x", "window_size", "palindrome_count", "palindrome_density", "GC_n")
-
-d.reduced = d[which(d$PTU %in% leading.region.PTUs), ]
-
-d.reduced$expected_density = theoreticalPalindromeExpectation(d.reduced$GC_n/d.reduced$window_size)
-
-d.reduced$diff_density = d.reduced$palindrome_density-d.reduced$expected_density
-
-# One line per plasmid
-d.reduced.median = d.reduced %>% group_by(PTU, x) %>% 
-  summarise(median=median(diff_density),
-            n=length(diff_density), 
-            median.original=median(palindrome_density),
-            GC=median(GC_n))
-
-# Maximum counts
-d.reduced.median.plot = d.reduced.median %>% 
-  group_by(PTU) %>%
-  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
-         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
-  ungroup() %>%  # Remove the grouping
-  select(-max_n)  # Optionally remove the max_n column if not needed
-# Plot the distribution
-p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median*1000))+
-  geom_hline(yintercept = 0, linetype='dashed')+
-  theme_bw()+
-  geom_path(aes(group=PTU, alpha=normalized_n))+
-  facet_wrap(~PTU, scales="free_x", nrow=3)+
-  ylab("Palindrome depletion per kb (5kb sliding window)")+
-  labs(alpha="% of plasmids within PTU")+
-  theme(panel.grid=element_blank())+  
-  scale_x_continuous(labels = function(x) x / 1000) +
-  xlab("Position on plasmid (kb)")+
-  theme(legend.position = c(0.8, 0.2))
-ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf',
-       width=10, height=6)
-# Plot only first 50kb
-ggsave(p.lines.expectation+scale_x_continuous(labels = function(x) x / 1000, limits=c(0,50000)), file='../results/6mer-deviation-expectation-0-50-kb.pdf',
-       width=10, height=6)
-
-```
-```{r}
-# Sliding window for GC content
-p.lines.GC = ggplot(d.reduced.median.plot, aes(x, GC/5000 * 100))+
-  geom_hline(yintercept = 0, linetype='dashed')+
-  theme_bw()+
-  geom_path(aes(group=PTU, alpha=normalized_n))+
-  facet_wrap(~PTU, scales="free_x", nrow=3)+
-  ylab("GC content (5kb sliding window)")+
-  labs(alpha="% of plasmids within PTU")+
-  theme(panel.grid=element_blank())+  
-  scale_x_continuous(labels = function(x) x / 1000) +
-  xlab("Position on plasmid (kb)")+
-  theme(legend.position = c(0.8, 0.2))
-ggsave(p.lines.GC, file='../results/GC-content.pdf',
-       width=10, height=6)
-
-# And for palindrome density (not expected - actual values)
-p.lines.density = ggplot(d.reduced.median.plot, aes(x, median.original*1000))+
-  geom_hline(yintercept = 0, linetype='dashed')+
-  theme_bw()+
-  geom_path(aes(group=PTU, alpha=normalized_n))+
-  facet_wrap(~PTU, scales="free_x", nrow=3)+
-  ylab("Palindrome density per kb (5kb sliding window)")+
-  labs(alpha="% of plasmids within PTU")+
-  theme(panel.grid=element_blank())+  
-  scale_x_continuous(labels = function(x) x / 1000) +
-  xlab("Position on plasmid (kb)")+
-  theme(legend.position = c(0.8, 0.2))
-ggsave(p.lines.density, file='../results/6mer-palindrome-density.pdf',
-       width=10, height=6)
-
-```
-
-### Figure 2
-
-Figure 2. RM target depletion in the leading region of conjugative plasmids.
-
-```{r}
-overall.median = d.reduced %>% filter(x<50000) %>%
-  group_by(x) %>%
-  summarise(y=median(diff_density),
-            n=length(diff_density))
-p.overall.average = ggplot(overall.median, aes(x, y*1000))+
-  geom_line()+
-  scale_y_continuous(breaks=seq(-10, 0, 2), limits = c(-9, 0))+
-  theme_bw()+
-  theme(panel.grid = element_blank())+
-  geom_hline(yintercept = 0, linetype='dashed')+
-  ylab("Palindrome depletion (per kb)")+
-    scale_x_continuous(labels = function(x) x / 1000) +
-  xlab("Position on plasmid (kb)")+
-  theme(axis.text=element_text(colour="black"))
-p.overall.average
-# Save
-ggsave(p.overall.average, file='../results/Figure-2-6mer-deviation-overall.pdf', 
-       width=3, height=3)
-```
-
-```{r}
-# Start from d.reduced, which contains per-plasmid results with 5000bp sliding window and step size of 500bp
-# Compare depletion in first 5kb with average for the rest of the plasmid
-diff.leading.region.average = d.reduced %>% 
-  group_by(plasmid, PTU) %>%
-  mutate(diff.count=palindrome_count[x=="0"]-mean(palindrome_count),
-         diff.density=diff_density[x=="0"]-mean(diff_density),
-         diff.GC=(GC_n[x=="0"]-mean(GC_n))/5000 * 100) %>%
-  filter(x=="0")
-
-# Order the PTUs by mean values of depletion for better visual display
-mean.values = diff.leading.region.average %>% group_by(PTU) %>%
-  summarise(mean=mean(diff.density)*1000)
-diff.leading.region.average$PTU = ordered(gsub("PTU-", "", diff.leading.region.average$PTU),
-                                          levels=gsub("PTU-", "", mean.values$PTU[order(mean.values$mean)]))
-p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
-  ggbeeswarm::geom_quasirandom(alpha=0.8)+
-  theme_bw()+
-  theme(panel.grid = element_blank())+
-  geom_hline(yintercept = 0)+
-  xlab("")+
-  theme(axis.text=element_text(colour="black", size=18))+
-  ylab("Palindrome depletion in leading region (per kb)")+
-  theme(axis.text.x=element_text(angle=45, hjust=1))+
-  ylab("")
-p.leading.region.diff.density
-# Based on this, remove two outliers
-p.leading.region.diff.density = p.leading.region.diff.density+
-  scale_y_continuous(breaks=seq(-8, 4, 2), limits=c(-8,5))
-ggsave(p.leading.region.diff.density, file='../results/Figure-2-6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
-       width=5.5, height=6)
-```
-
-### Figure 3
-
-Figure 3.
-GC-rich leading regions are more depleted in 6-bp palindromes.
-Values are calculated for the 5-10kb section of the leading region relative to the mean average value over the whole length of the plasmid.
-Spearman’s correlations are given as inset values.
-The relationship between relative GC content and palindrome depletion was still significant in a linear model controlling for PTU (depletion ~ GC*PTU).
-
-```{r}
-diff.PTU.averages = diff.leading.region.average %>% group_by(PTU) %>%
-  summarise(mean.GC=mean(diff.GC),
-            GC=mean(GC_n[x=="0"]/5000 * 100),
-            mean.palindrome=mean(palindrome_density),
-            mean.palindrome.diff=mean(diff.density) * 1000,
-            n=length(diff.GC))
-ggplot(diff.PTU.averages, aes(mean.GC, mean.palindrome.diff))+
-  geom_hline(yintercept = 0, linetype='dashed')+
-  ggrepel::geom_text_repel(aes(label=PTU))+
-    geom_point(aes(size=n), shape=21, colour="black", fill="grey")+
-  theme_bw()+
-  theme(panel.grid = element_blank())+
-  xlab("GC difference between first 5kb\nand plasmid average (%)")+
-  ylab("Additional palindrome depletion in leading region (per kb)")
-
-# Calculate actual GC and actual palindrome density
-# for a given x-value. I think 5kb is appropriate for leading region to 
-# get rid of start effects. 
-gc_data_frame = function(position_on_plasmid){
-  d.reduced %>% 
-  group_by(plasmid, PTU) %>%
-  mutate(GC=GC_n[x==position_on_plasmid]/5000 * 100,
-         diff.GC=(GC_n[x==position_on_plasmid]-mean(GC_n))/5000 * 100,
-         density=palindrome_density[x==position_on_plasmid]*1000,
-         density.depletion=diff_density[x==position_on_plasmid]*1000) %>%
-  filter(x==position_on_plasmid)
-}
-d.GC.palindrome = gc_data_frame(5000)
-ggplot(d.GC.palindrome, aes(diff.GC, density.depletion))+
-  geom_point()
-
-d.GC.palindrome.PTU = d.GC.palindrome %>% group_by(PTU) %>%
-  summarise(GC=mean(GC),
-            diff.GC=mean(diff.GC),
-            density=mean(density),
-            density.depletion=mean(density.depletion),
-            n=length(window_size))
-ggplot(d.GC.palindrome, aes(diff.GC, density.depletion, fill=PTU))+
-  geom_point(shape=21, colour="black")+
-  facet_wrap(~PTU)
-p.gc.correlation = ggplot(d.GC.palindrome, aes(diff.GC, density.depletion))+
-  geom_point()+
-  theme_bw()+
-  xlab("Relative GC content at 5-10kb (%)")+
-  ylab("Relative palindrome depletion at 5-10kb (per kb)")+
-    xlim(c(-15, 10))+
-  ylim(-13, 14)
-p.gc.correlation.ptu =  ggplot(d.GC.palindrome.PTU, aes(diff.GC, density.depletion))+
-  geom_point(aes(size=n, fill=PTU), shape=21, colour="black")+
-  theme_bw()+
-  ggrepel::geom_text_repel(aes(label=PTU), size=2, min.segment.length = 0)+
-  guides(fill=FALSE)+
-    xlab("Relative GC content (%)")+
-  ylab("Relative palindrome depletion (per kb)")
-pdf('../results/Figure-3-GC-correlation-depletion.pdf', width=8, height=4)
-cowplot::plot_grid(p.gc.correlation+ggtitle("(a)"), p.gc.correlation.ptu+ggtitle("(b)"))
-dev.off()
-cor.test(d.GC.palindrome$diff.GC, d.GC.palindrome$density.depletion, method="spearman")
-cor.test(d.GC.palindrome.PTU$diff.GC, d.GC.palindrome.PTU$density.depletion)
-
-summary(lm(density.depletion ~ diff.GC*PTU, data=d.GC.palindrome))
-
-```
-
-### Averaging over all plasmids
-
-Also do this for GC content.
-
-```{r}
-overall.median.GC = d.reduced %>% filter(x<50000) %>%
-  group_by(x) %>%
-  summarise(y=mean(GC_n),
-            n=length(GC_n))
-p.overall.average.GC = ggplot(overall.median.GC, aes(x, y/5000 * 100))+
-  geom_line()+
-  theme_bw()+
-  theme(panel.grid = element_blank())+
-  geom_hline(yintercept = 50, linetype='dashed')+
-  ylab("GC content (per kb)")+
-    scale_x_continuous(labels = function(x) x / 1000) +
-  xlab("Position on plasmid (kb)")+
-  theme(axis.text=element_text(colour="black"))
-p.overall.average.GC
-# Save
-ggsave(p.overall.average.GC, file='../results/GC-mean-overall.pdf', 
-       width=5, height=2)
-```
-
-```{r}
-lagging.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_lagging_10000_k6.csv', header=T)
-leading.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_leading_10000_k6.csv', header=T)
-merged.df = merge(leading.df, lagging.df, by="word")
-palindromes = tolower(read.csv('~/Downloads/trieste/k6.txt', header=F)$V1)
-ggplot(merged.df[which(merged.df$word %in% palindromes),], aes(score.x, score.y))+
-  geom_point()
-merged.df.palindromes = merged.df[which(merged.df$word %in% palindromes),]
-table(merged.df.palindromes$score.x<merged.df.palindromes$score.y)
-kmer_targets = tolower(read.csv('~/Downloads/trieste/PTU-LM_targets.txt', header=F)$V1)
-ggplot(merged.df[which(merged.df$word %in% kmer_targets),], aes(score.x, score.y))+
-  geom_point()
-merged.df.targets = merged.df[which(merged.df$word %in% kmer_targets),]
-table(merged.df.targets$score.x<merged.df.targets$score.y)
-
-```
-
-### Palindromes in leading vs. lagging
-
-```{r}
-scores_df = read.csv('../test_6mer_leading_lagging_scores.csv', header=T)
-scores_average <- scores_df %>%
-  rowwise() %>%
-  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
-  ungroup() %>%
-  select(kmer, score_avg)  
-# Order by average score
-scores_average$kmer = ordered(scores_average$kmer,
-                              levels=scores_average$kmer[order(scores_average$score_avg)])
-scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
-  coord_flip()
-
-# Same for counts
-counts_df = read.csv('../test_6mer_leading_lagging_counts.csv', header=T)
-counts_average <- counts_df %>%
-  rowwise() %>%
-  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
-  ungroup() %>%
-  select(kmer, count_avg)
-
-# Merge together
-merge_scores_counts = merge(counts_average, scores_average, by="kmer")
-# average score and average count are well-correlated
-ggplot(merge_scores_counts, aes(count_avg, score_avg))+geom_point()
-
-# Order by average score
-merge_scores_counts$kmer = ordered(merge_scores_counts$kmer,
-                              levels=scores_average$kmer[order(scores_average$score_avg)])
-scores.plot = ggplot(merge_scores_counts, aes(score_avg, kmer))+
-    geom_vline(xintercept = 0)+
-  geom_bar(stat="identity")+
-  geom_vline(xintercept = median(merge_scores_counts$score_avg), linetype='dashed', colour='red')+
-  theme_bw()+
-  theme(axis.text=element_text(colour="black", size=12),
-        axis.title.x=element_text(size=24),
-        axis.text.x=element_text(size=18))+
-  theme(axis.text.y=element_text(hjust=0))+
-  ylab("")+
-  xlab("Mean depletion score\n(leading vs. lagging, 10kb)")
-wilcox.test(merge_scores_counts$score_avg)
-ggsave(scores.plot, file='../results/6mer-palindromes-scores.pdf',
-       height=10, width=10)
-```
-
-### Figure 4
-
-Figure 4: Average depletion of within-range RM targets in the leading vs. lagging regions for conjugative PTUs. 
-```{r}
-
-results = NULL
-for (PTU in leading.region.PTUs){
-  for (k in c(5,6)){
-    scores_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_scores.csv'), header=T)
-    n_targets = nrow(scores_df)
-    scores_average <- scores_df %>%
-  rowwise() %>%
-  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
-  ungroup() %>%
-  select(kmer, score_avg)
-    counts_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_counts.csv'), header=T)
-    counts_average <- counts_df %>%
-  rowwise() %>%
-  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
-  ungroup() %>%
-  select(kmer, count_avg)
-    results = rbind(results, c(PTU, k, n_targets, median(scores_average$score_avg), median(counts_average$count_avg)))
-  }
-}
-
-results = data.frame(results)
-colnames(results) = c("PTU", "k", "n.targets", "median.score", "median.count")
-results$median.score= as.numeric(results$median.score)
-results$median.count = as.numeric(results$median.count)
-
-# Order by median score for k=6
-order.k.6.scores = results[which(results$k==6), "PTU"][order(results[which(results$k==6), "median.score"])]
-results$PTU = ordered(results$PTU,
-                      levels=rev(order.k.6.scores))
-# Add number of targets to each
-
-p.depletion.targets = ggplot(results, aes(median.score, as.factor(PTU)))+
-  geom_vline(xintercept = 0, size=1)+
-  geom_bar(stat="identity", position="dodge")+
-  facet_wrap(~k, scales="free_x")+
-  ylab("")+
-  xlab("Median depletion score of within-range RM targets\n(leading vs. lagging, 10kb)")+
-   theme_bw()+
-  theme(panel.grid = element_blank())+
-  theme(axis.text=element_text(colour="black", size=12),
-        axis.title.x=element_text(size=24),
-        axis.text.x=element_text(size=18))+
-  theme(axis.text.y=element_text(hjust=0))
-ggsave(p.depletion.targets, file='../results/Figure-4-figure-depletion-within-range-targets.pdf',
-       width=10, height=6)
-
-# Is there a correlation between k=5 and k=6 results?
-results.df = results %>% select(k, median.score, PTU) %>%
-  pivot_wider(
-    names_from = k,  # Use values of 'k' as new column names
-    values_from = median.score  # Use 'median.score' for the column values
-)
 ```
@@
 chisq.test(matrix(c(4, 37, 60, 4055 ), nrow=2))
-
 ```

### 2025-04-17 scripts/Paper-analysis.Rmd
more cleaning up
cleaning up

@@
 **Note** This relies on large downloaded files and is not intended to work out-of-the-box.
-For example, Roary gene presence/absence tables on my computer are stored in `/Users/Liam/Downloads/06_plaspan/`.
+For example, plasmid pangenome gene presence/absence tables are available on the associated Zenodo dataset as `plaspan.tar.gz`.
+This markdown document assumes they have been extracted to `/Users/Liam/Downloads/plaspan` (replace as appropriate).
 
@@
 }
-d = read.csv('../test_output_window5000_step500.csv', header=F, stringsAsFactors = F)
+d = read.csv('../results/output_window5000_step500.csv', header=F, stringsAsFactors = F)
 colnames(d) = c("PTU", "plasmid", "x", "window_size", "palindrome_count", "palindrome_density", "GC_n")
@@
   for (k in c(5,6)){
-    scores_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_scores.csv'), header=T)
+    scores_df = read.csv(paste0('../targets_leading_lagging/', PTU, '_k', k, '_scores.csv'), header=T)
     n_targets = nrow(scores_df)
@@
   select(kmer, score_avg)
-    counts_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_counts.csv'), header=T)
+    counts_df = read.csv(paste0('../targets_leading_lagging/', PTU, '_k', k, '_counts.csv'), header=T)
     counts_average <- counts_df %>%
@@
 ```{r}
-scores_df = read.csv('../test_6mer_leading_lagging_scores.csv', header=T)
+scores_df = read.csv('../results/6mer_leading_lagging_scores.csv', header=T)
 scores_average <- scores_df %>%
@@
 # Same for counts
-counts_df = read.csv('../test_6mer_leading_lagging_counts.csv', header=T)
+counts_df = read.csv('../6mer_leading_lagging_counts.csv', header=T)
 counts_average <- counts_df %>%
@@
 
-### Short palindromes (k=4, 6)
-
-```{r}
-# Pangenome statistics for PTUs
-#ptu = read.csv('../data/PTU-pangenome-stats.csv')
-
-# Read in RMES comparisons
-df_4 = read.csv('~/Downloads/trieste/test_k4.csv')
-df_6 = read.csv('~/Downloads/trieste/test_k6.csv')
-df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
-df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
-```
-
-Do the statistical test
-
-```{r}
-wilcox.test(df_4_filter$discrepancy_selected, df_4_filter$discrepancy_others, paired = TRUE)
-wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE)
-
-```
-
-### All RM targets (not included in main text)
+### All RM targets
 
@@
 # Read in results for RM targets
-df_4 = read.csv('../test_one_per_family_within_range_targets_k4.csv')
-df_5 = read.csv('../test_one_per_family_within_range_targets_k5.csv')
-df_6 = read.csv('../test_one_per_family_within_range_targets_k6.csv')
+df_4 = read.csv('../results/one_per_family_within_range_targets_k4.csv')
+df_5 = read.csv('../results/one_per_family_within_range_targets_k5.csv')
+df_6 = read.csv('../results/one_per_family_within_range_targets_k6.csv')
 
